---
title: "Previous-years comparison (v2) — Germany & federal states"
subtitle: "Is the 2026 surge in balcony-PV registrations really an outlier? When does it start?"
date: today
format:
  pdf:
    toc: true
    toc-depth: 3
    fig-width: 9
    fig-height: 5.5
    geometry:
      - margin=1.6cm
    fig-pos: "H"
    include-in-header:
      text: |
        \usepackage{float}
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    fig-width: 10
    fig-height: 6
    embed-resources: true
execute:
  warning: false
  message: false
  echo: false
---

> All charts overlay **2026 (red)** on **prior years (blue palette, light → dark)**. Data run through **2026-05-07**, so all year-on-year comparisons are restricted to the same Jan – early-May window.

# 1. Setup & event table

```{r setup}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here, arrow, dplyr, tidyr, lubridate, stringr, ggplot2, scales,
  data.table, purrr, forcats, ggrepel
)

theme_set(
  theme_minimal(base_size = 11) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title.position = "plot",
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(colour = "grey30"),
      legend.position = "top",
      legend.title = element_blank(),
      strip.text = element_text(face = "bold")
    )
)

# Iran war 2026 milestones (curated from the project's _milestones.qmd)
iran_milestones <- tibble::tribble(
  ~date,                  ~label,                                ~primary,
  as.Date("2026-01-08"),  "Iran protest crackdown peak",         FALSE,
  as.Date("2026-02-14"),  "US military buildup",                 FALSE,
  as.Date("2026-02-28"),  "War begins / Hormuz closed",          TRUE,
  as.Date("2026-03-18"),  "Major escalation",                    TRUE,
  as.Date("2026-04-08"),  "Temporary 2-week ceasefire",          TRUE
)

# Map any Date to a "fake" date in year 2000 so calendar position
# can be plotted on a single x-axis across years.
to_doy_date <- function(d) {
  d <- as.Date(d)
  as.Date(sprintf("2000-%02d-%02d", month(d), day(d)))
}

# Distinctive palette: sequential blues for 2022–2025, red for 2026.
year_palette <- c(
  "2022" = "#9ecae1",
  "2023" = "#4292c6",
  "2024" = "#08519c",
  "2025" = "#08306b",
  "2026" = "#d62728"
)

ma7 <- function(x) data.table::frollmean(x, 7, align = "right", na.rm = TRUE)
```

```{r load-data}
out_g <- here("4. datasets", "4.1. all germany", "1. data", "3")
germany_day   <- read_parquet(file.path(out_g, "germany_day.parquet"))
germany_week  <- read_parquet(file.path(out_g, "germany_week.parquet"))

out_s <- here("4. datasets", "4.2. federal state", "1. data", "3")
state_day  <- read_parquet(file.path(out_s, "state_day.parquet"))
state_week <- read_parquet(file.path(out_s, "state_week.parquet"))

last_day <- max(germany_day$date.registration.day)
last_doy <- to_doy_date(last_day)
```

```{r event-table}
iran_milestones |>
  mutate(
    date     = format(date, "%d %b %Y"),
    primary  = ifelse(primary, "yes", "no")
  ) |>
  rename(`Date` = date, `Event` = label, `Primary?` = primary) |>
  knitr::kable(caption = "Iran war 2026 — events used in this report.")
```

```{r milestone-helpers}
# Vertical lines + event labels (rotated) reused in calendar-overlay charts.
# Both lines and labels are black.
milestone_layers <- function(label_size = 3.0) {
  ml <- iran_milestones |>
    mutate(d = to_doy_date(date),
           lt = ifelse(primary, "solid", "dotted"))
  list(
    geom_vline(data = ml, aes(xintercept = d, linetype = lt),
               colour = "black", alpha = 0.7, linewidth = 0.45,
               show.legend = FALSE),
    scale_linetype_identity(),
    geom_text(data = ml,
              aes(x = d, y = Inf, label = label),
              inherit.aes = FALSE,
              hjust = 1.02, vjust = -0.4,
              size = label_size, colour = "black", angle = 90,
              fontface = "bold")
  )
}
```

---

# 2. Figure 1 — local view around each war date (anticipation diagnostic)

> Outcome: **daily new balcony-PV registrations per 100k** in a **±21-day window** around each event (day 0 = event). Grey bars = raw daily counts; red line = centred 7-day moving average ($\bar n_t^{(7)} = \frac{1}{7}\sum_{k=-3}^{3} n_{t+k}$). The vertical dashed line marks day 0.
> Reading guide: if the red line is already rising in the negative-day region (left of the dashed line), that is an **anticipation** signal.

```{r fig1-local-events, fig.height=8, fig.width=11}
window_days <- 21

daily_balcony <- germany_day |>
  filter(solar_category == "Balcony PVs") |>
  select(date = date.registration.day, n_per100k)

event_windows <- iran_milestones |>
  rowwise() |>
  mutate(window = list(seq(date - window_days, date + window_days, by = "day"))) |>
  ungroup() |>
  tidyr::unnest(window) |>
  mutate(rel_day = as.integer(window - date)) |>
  left_join(daily_balcony, by = c("window" = "date")) |>
  mutate(
    n_per100k = tidyr::replace_na(n_per100k, 0),
    event_label = sprintf("%s — %s",
                          format(date, "%d %b %Y"),
                          label)
  ) |>
  group_by(event_label, date) |>
  arrange(rel_day, .by_group = TRUE) |>
  mutate(ma7c = data.table::frollmean(n_per100k, 7, align = "center", na.rm = TRUE)) |>
  ungroup() |>
  mutate(event_label = forcats::fct_reorder(event_label, date))

ggplot(event_windows, aes(x = rel_day)) +
  geom_col(aes(y = n_per100k), fill = "grey55", alpha = 0.7, width = 1) +
  geom_line(aes(y = ma7c), colour = "#d62728", linewidth = 1.0) +
  geom_vline(xintercept = 0, colour = "black",
             linetype = "dashed", linewidth = 0.5) +
  facet_wrap(~ event_label, ncol = 2, scales = "free_y") +
  scale_x_continuous(breaks = seq(-window_days, window_days, by = 7)) +
  labs(
    title = "Figure 1 — Daily balcony-PV registrations around each war milestone",
    subtitle = sprintf("Germany. ±%d days around event (day 0 = event). Grey bars = raw daily; red line = centred 7-day MA.",
                       window_days),
    x = "Days relative to event",
    y = "Balcony PVs / 100k (per day)"
  )
```

---

# 3. Figure 2 — Germany year-to-date totals (Jan – early May)

> Outcome: number of new registrations from 1 January to `r format(last_day, "%d %B")` of each year, per 100k inhabitants. Like-for-like window across years.

```{r fig2-ytd-bars, fig.height=4.2}
ytd <- germany_day |>
  mutate(year = year(date.registration.day),
         mmdd = format(date.registration.day, "%m-%d")) |>
  filter(year >= 2022, mmdd <= format(last_day, "%m-%d")) |>
  group_by(solar_category, year) |>
  summarise(n = sum(n), n_per100k = sum(n_per100k), .groups = "drop")

ggplot(ytd, aes(x = factor(year), y = n_per100k,
                fill = factor(year) == "2026")) +
  geom_col(width = 0.7) +
  geom_text(aes(label = scales::comma(round(n_per100k, 1))),
            vjust = -0.4, size = 3.2) +
  facet_wrap(~ solar_category, scales = "free_y") +
  scale_fill_manual(values = c(`FALSE` = "grey60", `TRUE` = "#d62728"),
                    guide = "none") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(
    title = sprintf("Figure 2 — Year-to-date registrations through %s, by year",
                    format(last_day, "%d %B")),
    subtitle = "Outcome: number of new registrations per 100k inhabitants (Germany)",
    x = NULL, y = "Registrations / 100k (Jan–early May)"
  )
```

---

# 4. Daily registrations, calendar overlay across years

## Figure 3.1 — Balcony PVs, 7-day moving average

> Outcome: 7-day moving average of daily new balcony-PV registrations, per 100k.
> Formula: $\bar n_t^{(7)} = \frac{1}{7}\sum_{k=0}^{6} n_{t-k}$ (right-aligned).
> Vertical lines = Iran-war milestones, labelled.

```{r fig3-1-daily-balcony, fig.height=6}
df_day <- germany_day |>
  filter(solar_category == "Balcony PVs",
         year(date.registration.day) >= 2022) |>
  arrange(date.registration.day) |>
  mutate(year = factor(year(date.registration.day))) |>
  group_by(year) |>
  mutate(n_per100k_ma7 = ma7(n_per100k)) |>
  ungroup() |>
  mutate(doy_date = to_doy_date(date.registration.day)) |>
  filter(doy_date <= last_doy)

# Last in-window point per year, for the end-of-line year label
end_labels_31 <- df_day |>
  filter(!is.na(n_per100k_ma7)) |>
  group_by(year) |>
  slice_max(doy_date, n = 1) |>
  ungroup()

ggplot(df_day, aes(x = doy_date, y = n_per100k_ma7,
                   colour = year, linewidth = year == "2026")) +
  geom_line() +
  milestone_layers() +
  ggrepel::geom_text_repel(
    data = end_labels_31,
    aes(label = year, colour = year),
    direction = "y", hjust = 0, nudge_x = 4,
    segment.size = 0.2, segment.alpha = 0.5,
    box.padding = 0.15, point.padding = 0.1,
    size = 3.4, fontface = "bold", show.legend = FALSE
  ) +
  scale_colour_manual(values = year_palette, guide = "none") +
  scale_linewidth_manual(values = c(`FALSE` = 0.95, `TRUE` = 1.6),
                         guide = "none") +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%d %b",
               limits = c(as.Date("2000-01-01"), last_doy + 12),
               expand = expansion(mult = c(0.01, 0.02))) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(
    title = "Figure 3.1 — Balcony PVs: daily new registrations (7-day MA), Jan–early May overlay",
    subtitle = "Germany. 2026 (red) vs 2022–2025 (blue palette, light → dark). Per 100k inhabitants.",
    x = NULL, y = "Balcony PVs / 100k (7-day MA)"
  )
```

## Figure 3.2 — Non-balcony PVs, 7-day moving average (control)

```{r fig3-2-daily-nonbalcony, fig.height=6}
df_day_nb <- germany_day |>
  filter(solar_category == "Non-balcony PVs",
         year(date.registration.day) >= 2022) |>
  arrange(date.registration.day) |>
  mutate(year = factor(year(date.registration.day))) |>
  group_by(year) |>
  mutate(n_per100k_ma7 = ma7(n_per100k)) |>
  ungroup() |>
  mutate(doy_date = to_doy_date(date.registration.day)) |>
  filter(doy_date <= last_doy)

end_labels_32 <- df_day_nb |>
  filter(!is.na(n_per100k_ma7)) |>
  group_by(year) |>
  slice_max(doy_date, n = 1) |>
  ungroup()

ggplot(df_day_nb, aes(x = doy_date, y = n_per100k_ma7,
                      colour = year, linewidth = year == "2026")) +
  geom_line() +
  milestone_layers() +
  ggrepel::geom_text_repel(
    data = end_labels_32,
    aes(label = year, colour = year),
    direction = "y", hjust = 0, nudge_x = 4,
    segment.size = 0.2, segment.alpha = 0.5,
    box.padding = 0.15, point.padding = 0.1,
    size = 3.4, fontface = "bold", show.legend = FALSE
  ) +
  scale_colour_manual(values = year_palette, guide = "none") +
  scale_linewidth_manual(values = c(`FALSE` = 0.95, `TRUE` = 1.6),
                         guide = "none") +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%d %b",
               limits = c(as.Date("2000-01-01"), last_doy + 12),
               expand = expansion(mult = c(0.01, 0.02))) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(
    title = "Figure 3.2 — Non-balcony PVs: daily new registrations (7-day MA), Jan–early May overlay",
    subtitle = "Germany. Control series: shock should be much weaker or absent here.",
    x = NULL, y = "Non-balcony PVs / 100k (7-day MA)"
  )
```

## Figure 3.3 — Balcony PVs, raw daily registrations (no smoothing)

> Same overlay as 3.1 but using the **raw daily count** instead of the 7-day MA. Noisier but honest — every spike is a real day.

```{r fig3-3-daily-balcony-raw, fig.height=6}
end_labels_33 <- df_day |>
  filter(!is.na(n_per100k)) |>
  group_by(year) |>
  slice_max(doy_date, n = 1) |>
  ungroup()

ggplot(df_day, aes(x = doy_date, y = n_per100k,
                   colour = year, linewidth = year == "2026",
                   alpha = year == "2026")) +
  geom_line() +
  milestone_layers() +
  ggrepel::geom_text_repel(
    data = end_labels_33,
    aes(label = year, colour = year),
    direction = "y", hjust = 0, nudge_x = 4,
    segment.size = 0.2, segment.alpha = 0.5,
    box.padding = 0.15, point.padding = 0.1,
    size = 3.4, fontface = "bold", alpha = 1, show.legend = FALSE
  ) +
  scale_colour_manual(values = year_palette, guide = "none") +
  scale_linewidth_manual(values = c(`FALSE` = 0.55, `TRUE` = 1.0),
                         guide = "none") +
  scale_alpha_manual(values = c(`FALSE` = 0.55, `TRUE` = 1.0),
                     guide = "none") +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%d %b",
               limits = c(as.Date("2000-01-01"), last_doy + 12),
               expand = expansion(mult = c(0.01, 0.02))) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(
    title = "Figure 3.3 — Balcony PVs: raw daily new registrations, Jan–early May overlay",
    subtitle = "Germany. Raw daily counts, no smoothing. Per 100k inhabitants.",
    x = NULL, y = "Balcony PVs / 100k (per day)"
  )
```

---

# 5. Figure 6 — Germany cumulative year-to-date

> Outcome: cumulative number of new balcony-PV registrations per 100k, reset on 1 January of each year. The day where the 2026 line breaks away from the prior-year lines marks the start of the cumulative deviation.

```{r fig6-cumulative, fig.height=6}
df_cum <- germany_day |>
  filter(solar_category == "Balcony PVs",
         year(date.registration.day) >= 2022) |>
  arrange(date.registration.day) |>
  mutate(year = factor(year(date.registration.day))) |>
  group_by(year) |>
  mutate(cum_per100k_ytd = cumsum(n_per100k)) |>
  ungroup() |>
  mutate(doy_date = to_doy_date(date.registration.day)) |>
  filter(doy_date <= last_doy)

ggplot(df_cum, aes(x = doy_date, y = cum_per100k_ytd,
                   colour = year, linewidth = year == "2026")) +
  geom_line() +
  milestone_layers() +
  scale_colour_manual(values = year_palette) +
  scale_linewidth_manual(values = c(`FALSE` = 0.95, `TRUE` = 1.6),
                         guide = "none") +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%d %b",
               limits = c(as.Date("2000-01-01"), last_doy),
               expand = expansion(mult = c(0.01, 0.02))) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(
    title = "Figure 6 — Balcony PVs: cumulative year-to-date registrations",
    subtitle = "Germany. Counter resets on 1 January each year. Per 100k inhabitants.",
    x = NULL, y = "Cumulative balcony PVs / 100k (YTD)"
  )
```

---

# 6. Figure 7 — Germany weekly view (less noise)

> Outcome: number of weekly new balcony-PV registrations per 100k. Smoother than daily; easier to read off the post-war jump.

```{r fig7-weekly, fig.height=6}
df_w <- germany_week |>
  filter(solar_category == "Balcony PVs",
         year(date.registration.week) >= 2022) |>
  mutate(year     = factor(year(date.registration.week)),
         doy_date = to_doy_date(date.registration.week)) |>
  filter(doy_date <= last_doy)

ggplot(df_w, aes(x = doy_date, y = n_per100k,
                 colour = year, linewidth = year == "2026")) +
  geom_line() +
  geom_point(data = subset(df_w, year == "2026"), size = 1.3) +
  milestone_layers() +
  scale_colour_manual(values = year_palette) +
  scale_linewidth_manual(values = c(`FALSE` = 0.95, `TRUE` = 1.6),
                         guide = "none") +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%d %b",
               limits = c(as.Date("2000-01-01"), last_doy),
               expand = expansion(mult = c(0.01, 0.02))) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(
    title = "Figure 7 — Balcony PVs: weekly new registrations, Jan–early May overlay",
    subtitle = "Germany. Per 100k inhabitants.",
    x = NULL, y = "Balcony PVs / 100k (per week)"
  )
```

---

# 7. Federal-state heterogeneity

## Figure 8.1 — Weekly registrations per 100k, by federal state

> Same calendar overlay as Figure 7 but faceted across the 16 federal states. Larger linewidths to distinguish each year.

```{r fig81-state-weekly, fig.height=11, fig.width=12}
df_sw <- state_week |>
  filter(solar_category == "Balcony PVs",
         year(date.registration.week) >= 2022) |>
  mutate(year     = factor(year(date.registration.week)),
         doy_date = to_doy_date(date.registration.week)) |>
  filter(doy_date <= last_doy)

ggplot(df_sw, aes(x = doy_date, y = n_per100k,
                  colour = year, linewidth = year == "2026")) +
  geom_line() +
  geom_vline(data = iran_milestones |> filter(primary) |>
               mutate(d = to_doy_date(date)),
             aes(xintercept = d), colour = "black",
             alpha = 0.7, linetype = "solid", linewidth = 0.4) +
  facet_wrap(~ FederalState.names, ncol = 4, scales = "free_y") +
  scale_colour_manual(values = year_palette) +
  scale_linewidth_manual(values = c(`FALSE` = 0.8, `TRUE` = 1.3),
                         guide = "none") +
  scale_x_date(date_breaks = "2 months", date_labels = "%b",
               limits = c(as.Date("2000-01-01"), last_doy)) +
  theme(axis.text.x = element_text(angle = 0, size = 7),
        axis.text.y = element_text(size = 7),
        strip.text = element_text(size = 8)) +
  labs(
    title = "Figure 8.1 — Balcony PVs: weekly registrations per 100k, by federal state",
    subtitle = "2026 (red) vs 2022–2025. Black verticals = primary events: war start (28 Feb), major escalation (18 Mar), ceasefire (8 Apr).",
    x = NULL, y = "Balcony PVs / 100k (per week)"
  )
```

## Figure 8.3 — Year-to-date totals by federal state and year

```{r fig83-state-ytd-bars, fig.height=4.6, fig.width=12}
ytd_state <- state_day |>
  filter(solar_category == "Balcony PVs") |>
  mutate(year = year(date.registration.day),
         mmdd = format(date.registration.day, "%m-%d")) |>
  filter(year >= 2022, mmdd <= format(last_day, "%m-%d")) |>
  group_by(FederalState.names, year) |>
  summarise(n_per100k = sum(n_per100k), .groups = "drop")

order_fs <- ytd_state |>
  filter(year == 2026) |>
  arrange(desc(n_per100k)) |>
  pull(FederalState.names)

ytd_state |>
  mutate(FederalState.names = factor(FederalState.names, levels = order_fs)) |>
  ggplot(aes(x = FederalState.names, y = n_per100k,
             fill = factor(year))) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  scale_fill_manual(values = year_palette) +
  theme(axis.text.x = element_text(angle = 35, hjust = 1)) +
  labs(
    title = sprintf("Figure 8.3 — Balcony PVs: Jan–%s YTD registrations per 100k, by federal state and year",
                    format(last_day, "%d %b")),
    subtitle = "States ordered by 2026 intensity (highest left).",
    x = NULL, y = "Balcony PVs / 100k (Jan–early May)"
  )
```

