---
title: 'Dataset for all municipalities'
format: html
---

> Here I want to create the dataset for the municipalities at daily frequency.

**Shiny app I want to create**:
- I want to create a shiny app where on the left there is a map of all municipalities (with federla states bounderies) and the outcome variable to consider must be the registrations (population adjusted) cumulative, from the beginning of the sample to the end.
- on the right there must a chart of the daily registration in grey + MA
Once you seelct on the map one munciipality, it dispays the chart of that specific munciiaplity (from beginning janaury 2026 up to most recent data) of the daily registrations (popuation corrected) on the backgroung in grey + moving average (with a slide to deicde the window)
As default munciiaplity to show at the beginning, even before seelcting a specific municipality, show the plot for Mannheim municipality
- in the plot add the vertcial line sin black for the iran war relevant dates (max 3 dates, no more)
- create an option where you can decide whether to focs on balcoy PVs or non-balcoy PVs

Here you have the datasets:
  1. dataset for munciipality bounderies: *8. shiny app\8.1. shiny apps\3\municipalities.parquet*
  2. dataset for federal states bounderies: *8. shiny app\8.1. shiny apps\3\federalState.parquet*
  3. the dataset where you cna find the registration is here: *8. shiny app\8.1. shiny apps\3\mun_day.parquet*

Here (*8. shiny app\8.1. shiny apps\3\1. dataset.qmd*) you can find the file i used to create the dataset for the balcony PVs registrations


**Claude chat**: SHINY APP (3) (and maybe SHINY APP (2))

**Libraries**:

```{r}
pacman::p_load(
  here,        # here()
  arrow,       # open_dataset(), read_parquet(), write_parquet(), collect()
  sfarrow,     # st_read_parquet()
  sf,          # st_drop_geometry()
  dplyr,       # rename, mutate, case_when, if_else, left_join, relocate,
               # filter, select, group_by, summarise, arrange, pull, ungroup, n
  tibble,      # tibble()
  tidyr,       # complete()
  lubridate,   # floor_date()
  stringr,     # str_pad()
  restatis,    # gen_logincheck(), gen_table()
  shiny        # shinyApp(), fluidPage(), selectizeInput(), updateSelectizeInput(), renderPrint(), verbatimTextOutput(), req()
)
```


**Frequency**:
- daily registrations
- I want to focus only on the Iran war (so you can discard all the observations before that)

# 1. Dataset
> Here I want to generate the code from zero.

**Dataset (not population adjusted)**:


```{r}
data <- open_dataset(here("01_data/1.2. May 7 2026/EinheitenSolar.parquet"))


# Step 1: rename German → English (explicit mapping, column order differs from old dataset)
data <- data |>
  rename(
    UnitMastrNumber                          = EinheitMastrNummer,
    DateLastUpdate                           = DatumLetzteAktualisierung,
    LocationMastrNumber                      = LokationMaStRNummer,
    GridOperatorVerificationStatus           = NetzbetreiberpruefungStatus,
    GridOperatorVerificationDate             = NetzbetreiberpruefungDatum,
    PlantOperatorMastrNumber                 = AnlagenbetreiberMastrNummer,
    Country                                  = Land,
    FederalState                             = Bundesland,
    District                                 = Landkreis,
    Municipality                             = Gemeinde,
    MunicipalityCode                         = Gemeindeschluessel,
    PostalCode                               = Postleitzahl,
    City                                     = Ort,
    RegistrationDate                         = Registrierungsdatum,
    CommissioningDate                        = Inbetriebnahmedatum,
    UnitSystemStatus                         = EinheitSystemstatus,
    UnitOperationalStatus                    = EinheitBetriebsstatus,
    NotPresentInMigratedUnits                = NichtVorhandenInMigriertenEinheiten,
    PowerGenerationUnitName                  = NameStromerzeugungseinheit,
    Weic_NotAvailable                        = Weic_nv,
    PowerPlantNumber_NotAvailable            = Kraftwerksnummer_nv,
    EnergySource                             = Energietraeger,
    GrossCapacity                            = Bruttoleistung,
    NetNominalCapacity                       = Nettonennleistung,
    RemoteControllabilityGridOperator        = FernsteuerbarkeitNb,
    FeedInType                               = Einspeisungsart,
    AssignedActivePowerInverter              = ZugeordneteWirkleistungWechselrichter,
    NumberOfModules                          = AnzahlModule,
    TypeOfSolarSystem                        = ArtDerSolaranlage,
    PowerLimitation                          = Leistungsbegrenzung,
    UniformOrientationAndTiltAngle           = EinheitlicheAusrichtungUndNeigungswinkel,
    MainOrientation                          = Hauptausrichtung,
    MainOrientationTiltAngle                 = HauptausrichtungNeigungswinkel,
    UsageArea                                = Nutzungsbereich,
    EegMastrNumber                           = EegMaStRNummer,
    StorageAtSameLocation                    = SpeicherAmGleichenOrt,
    SecondaryOrientation                     = Nebenausrichtung,
    SecondaryOrientationTiltAngle            = NebenausrichtungNeigungswinkel,
    Street                                   = Strasse,
    StreetNotFound                           = StrasseNichtGefunden,
    HouseNumber_NotAvailable                 = Hausnummer_nv,
    HouseNumberNotFound                      = HausnummerNichtGefunden,
    Longitude                                = Laengengrad,
    Latitude                                 = Breitengrad,
    DateStartTemporaryDecommissioning        = DatumBeginnVoruebergehendeStilllegung,
    HouseNumber                              = Hausnummer,
    AddressAddition                          = Adresszusatz,
    RemoteControllabilityDirectMarketer      = FernsteuerbarkeitDv,
    PreviousPlantOperatorMastrNumber         = AltAnlagenbetreiberMastrNummer,  # new variable
    DateOfOperatorChange                     = DatumDesBetreiberwechsels,
    DateRegistrationOfOperatorChange         = DatumRegistrierungDesBetreiberwechsels,
    CadastralDistrict                        = Gemarkung,
    FieldParcelNumbers                       = FlurFlurstuecknummern,
    ConnectionToExtraHighOrHighVoltage       = AnschlussAnHoechstOderHochSpannung,
    DatePermanentDecommissioning             = DatumEndgueltigeStilllegung,
    SizeOfUsedAreaInHectares                 = GroesseDerInAnspruchGenommenenFlaecheInHektar,
    PreviousUsageTypeOfArea                  = VorherigerNutzungsartenbereichDerFlaeche,
    PredominantUsageTypeBeforeSolarConstruction = UeberwiegendeNutzungsartDerFlaecheVorErrichtung,
    NameOfSolarPark                          = NameDesSolarparks,
    AdditionalFeaturesOfAreaAndCurrentUsage  = ZusaetzlicheMerkmaleDerFlaecheUndDerAktuellenFlaechennutzung,
    GenMastrNumber                           = GenMastrNummer,
    CommissioningDateAtCurrentLocation       = InbetriebnahmedatumAmAktuellenStandort,
    DispatchResponsible                      = Einsatzverantwortlicher,
    CitizenEnergy                            = Buergerenergie,
    DateResumptionOfOperation                = DatumWiederaufnahmeBetrieb,
    PlannedCommissioningDate                 = GeplantesInbetriebnahmedatum,
    ClearanceHeight                          = LichteHoehe,
    PowerPlantNumber                         = Kraftwerksnummer
  )

# Step 2: convert numeric columns (XML parsing stores everything as character)
data <- data |>
  mutate(
    TypeOfSolarSystem           = as.integer(TypeOfSolarSystem),
    Country                     = as.integer(Country),
    FederalState                = as.integer(FederalState),
    GrossCapacity               = as.numeric(GrossCapacity),
    NetNominalCapacity          = as.numeric(NetNominalCapacity),
    AssignedActivePowerInverter = as.numeric(AssignedActivePowerInverter),
    NumberOfModules             = as.integer(NumberOfModules)
  )

# Step 3: lookup tables
C1 <- data.frame(
  TypeOfSolarSystem.names = c(
    "Structural_facilities", "Other_Structural_facilities", "Open_space",
    "Bodies_water", "Large_parking_lot", "Plug_in_sola_system", "Onshore&Offshore_wind_power"
  ),
  TypeOfSolarSystem = as.integer(c(853, 2484, 852, 3002, 3058, 2961, NA))
)

C2 <- tibble(
  Country = as.integer(c(84, 231, 90, 66, 198, 206, 106)),
  Country.sn = c("Germany", "Switzerland", "Denmark", "Belgium", "Netherlands", "Austria", "France")
)

C3 <- data.frame(
  FederalState = as.integer(c(
    1409, 1402, 1400, 1403, 1412, 1408, 1410, 1407,
    1414, 1411, 1413, 1405, 1415, 1406, 1401, 1404
  )),
  FederalState.names = c(
    "North Rhine-Westphalia", "Baden-Württemberg", "Brandenburg", "Bavaria",
    "Saarland", "Lower Saxony", "Rhineland-Palatinate", "Mecklenburg-Western Pomerania",
    "Saxony-Anhalt", "Schleswig-Holstein", "Saxony", "Hesse",
    "Thuringia", "Hamburg", "Berlin", "Bremen"
  ),
  FederalStateCode = c(
    "DEA", "DE1", "DE4", "DE2",
    "DEC", "DE9", "DEB", "DEM",
    "DEE", "DEF", "DED", "DE7",
    "DEG", "DE6", "DE3", "DE5"
  )
)

# district code from map (no geometry needed)
districts_map <- st_read_parquet(here("ignore/3. data/2. maps/districts.parquet")) |>
  st_drop_geometry() |>
  select(District, DistrictCode)

# Step 4: join lookups
data <- data |>
  left_join(C1, by = "TypeOfSolarSystem") |>
  relocate(TypeOfSolarSystem.names, .after = TypeOfSolarSystem) |>
  left_join(C2, by = "Country") |>
  relocate(Country.sn, .after = Country) |>
  left_join(C3, by = "FederalState") |>
  relocate(FederalState.names, .after = FederalState) |>
  relocate(FederalStateCode, .after = FederalState.names) |>
  left_join(districts_map, by = "District") |>
  relocate(DistrictCode, .after = District)

# Step 5: solar_category
reform_date <- as.Date("2024-05-16")

data <- data |>
  mutate(
    RegistrationDate  = as.Date(RegistrationDate),
    CommissioningDate = as.Date(CommissioningDate),
    solar_category = case_when(
      TypeOfSolarSystem == 2961 & RegistrationDate <  reform_date & AssignedActivePowerInverter <= 0.6 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 & RegistrationDate >= reform_date & AssignedActivePowerInverter <= 0.8 & GrossCapacity <= 2.0 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 ~ "Plug_in_sola_system - Non-Balcony",
      TRUE ~ TypeOfSolarSystem.names
    )
  )

# Step 6: filter Germany
data <- data |>
  filter(Country.sn == "Germany")

# Step 7: date variables (week, month, quarter, year)
data <- data |>
  mutate(
    date.registration.day     = RegistrationDate,
    date.registration.week    = floor_date(RegistrationDate,  "week"),
    date.registration.month   = floor_date(RegistrationDate,  "month"),
    date.registration.quarter = floor_date(RegistrationDate,  "quarter"),
    date.registration.year    = floor_date(RegistrationDate,  "year"),
    date.commission.day       = CommissioningDate,
    date.commission.week      = floor_date(CommissioningDate, "week"),
    date.commission.month     = floor_date(CommissioningDate, "month"),
    date.commission.quarter   = floor_date(CommissioningDate, "quarter"),
    date.commission.year      = floor_date(CommissioningDate, "year")
  )

# Step 8: fix MunicipalityCode
data <- data |>
  mutate(MunicipalityCode = str_pad(as.character(MunicipalityCode), 8, pad = "0"))

# Collapse every category other than `Balcony PVs` into a single `Non-balcony PVs` group:


data <- data |>
  mutate(solar_category = if_else(solar_category == "Plug_in_sola_system - Balcony",
                                  "Balcony PVs",
                                  "Non-balcony PVs"))

# I want to focus only on the Iran war (so you can discard all the observations before that):
sample_start   <- as.Date("2026-01-01")

data <- data |>
  filter(RegistrationDate >= sample_start) |>
  collect()

```

Frequencies (only daily): 

```{r}
mun_day <- data |>
  group_by(MunicipalityCode, Municipality, solar_category, date.registration.day) |>
  summarise(n = n(), .groups = "drop") |>
  complete(nesting(MunicipalityCode, Municipality), solar_category,
           date.registration.day = seq(min(date.registration.day), max(date.registration.day), by = "day"),
           fill = list(n = 0)) |>
  arrange(MunicipalityCode, solar_category, date.registration.day) |>
  group_by(MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```


Frequencies (weekly, monthly, quarterly, yearly):
```{r}
mun_week <- data |>
  group_by(MunicipalityCode, Municipality, solar_category, date.registration.week) |>
  summarise(n = n(), .groups = "drop") |>
  complete(nesting(MunicipalityCode, Municipality), solar_category,
           date.registration.week = seq(min(date.registration.week), max(date.registration.week), by = "week"),
           fill = list(n = 0)) |>
  arrange(MunicipalityCode, solar_category, date.registration.week) |>
  group_by(MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_month <- data |>
  group_by(MunicipalityCode, Municipality, solar_category, date.registration.month) |>
  summarise(n = n(), .groups = "drop") |>
  complete(nesting(MunicipalityCode, Municipality), solar_category,
           date.registration.month = seq(min(date.registration.month), max(date.registration.month), by = "month"),
           fill = list(n = 0)) |>
  arrange(MunicipalityCode, solar_category, date.registration.month) |>
  group_by(MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_quarter <- data |>
  group_by(MunicipalityCode, Municipality, solar_category, date.registration.quarter) |>
  summarise(n = n(), .groups = "drop") |>
  complete(nesting(MunicipalityCode, Municipality), solar_category,
           date.registration.quarter = seq(min(date.registration.quarter), max(date.registration.quarter), by = "quarter"),
           fill = list(n = 0)) |>
  arrange(MunicipalityCode, solar_category, date.registration.quarter) |>
  group_by(MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_year <- data |>
  group_by(MunicipalityCode, Municipality, solar_category, date.registration.year) |>
  summarise(n = n(), .groups = "drop") |>
  complete(nesting(MunicipalityCode, Municipality), solar_category,
           date.registration.year = seq(min(date.registration.year), max(date.registration.year), by = "year"),
           fill = list(n = 0)) |>
  arrange(MunicipalityCode, solar_category, date.registration.year) |>
  group_by(MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```


**Population**:

```{r}
pop.muni.2023 <- st_read_parquet(
  here("6. Maps", "6.2. map for 2023", "2. final dataset", "municipalities.parquet")
) |>
  st_drop_geometry() |>
  select(MunicipalityCode, Municipality, population = Population2023)
```


**Merging**:
```{r}
pop_lookup <- pop.muni.2023 |>
  select(MunicipalityCode, population)

mun_day <- mun_day |>
  left_join(pop_lookup, by = "MunicipalityCode") |>
  mutate(
    n_per100k            = (n            / population) * 100000,
    n_cumulative_per100k = (n_cumulative / population) * 100000
  )
```

Here I want you to select only those municipalities that are in the file: *6. Maps\6.2. map for 2023\2. final dataset\municipalities.parquet* (since only these municiplaities will be displayed in the map):

```{r}
# pop.muni.2023 was loaded straight from that file, so its codes ARE the map set
map_codes <- pop.muni.2023$MunicipalityCode

mun_day     <- mun_day     |> filter(MunicipalityCode %in% map_codes)
mun_week    <- mun_week    |> filter(MunicipalityCode %in% map_codes)
mun_month   <- mun_month   |> filter(MunicipalityCode %in% map_codes)
mun_quarter <- mun_quarter |> filter(MunicipalityCode %in% map_codes)
mun_year    <- mun_year    |> filter(MunicipalityCode %in% map_codes)
```


Other frequencies:
```{r}
mun_week <- mun_week |>
  left_join(pop_lookup, by = "MunicipalityCode") |>
  mutate(
    n_per100k            = (n            / population) * 100000,
    n_cumulative_per100k = (n_cumulative / population) * 100000
  )

mun_month <- mun_month |>
  left_join(pop_lookup, by = "MunicipalityCode") |>
  mutate(
    n_per100k            = (n            / population) * 100000,
    n_cumulative_per100k = (n_cumulative / population) * 100000
  )

mun_quarter <- mun_quarter |>
  left_join(pop_lookup, by = "MunicipalityCode") |>
  mutate(
    n_per100k            = (n            / population) * 100000,
    n_cumulative_per100k = (n_cumulative / population) * 100000
  )

mun_year <- mun_year |>
  left_join(pop_lookup, by = "MunicipalityCode") |>
  mutate(
    n_per100k            = (n            / population) * 100000,
    n_cumulative_per100k = (n_cumulative / population) * 100000
  )
```

**Save**:

```{r}
out_dir <- here("8. shiny app", "8.1. shiny apps", "2")

write_parquet(mun_day,     file.path(out_dir, "mun_day.parquet"))
write_parquet(mun_week,    file.path(out_dir, "mun_week.parquet"))
write_parquet(mun_month,   file.path(out_dir, "mun_month.parquet"))
write_parquet(mun_quarter, file.path(out_dir, "mun_quarter.parquet"))
write_parquet(mun_year,    file.path(out_dir, "mun_year.parquet"))
```

**Load**:

```{r}
out_dir <- here("8. shiny app", "8.1. shiny apps", "2")

mun_day     <- read_parquet(file.path(out_dir, "mun_day.parquet"))
mun_week    <- read_parquet(file.path(out_dir, "mun_week.parquet"))
mun_month   <- read_parquet(file.path(out_dir, "mun_month.parquet"))
mun_quarter <- read_parquet(file.path(out_dir, "mun_quarter.parquet"))
mun_year    <- read_parquet(file.path(out_dir, "mun_year.parquet"))
```

