---
title: 'Municipalities geometry'
format: html
---



> Here I want to create the daatset with geometry for municiplaities using year 2023.
Link: https://ec.europa.eu/eurostat/web/gisco/geodata/statistical-units/local-administrative-units
Options:
- File Format: GeoPackage
- Coordinate reference system: EPSG 4326

**Libraries**:

```{r}
pacman::p_load(here, sf, sfarrow, dplyr, leaflet, restatis)
```


::: {.callout}
**What I want**:
I need a dataset with:

- municiplaity name: ciioaoa
- municipality code
- geometry
- some additional variables (that I will remove later maybe)

Must be only for Germany, not other countries.





# 1. load the dataset
> Here I want to load the dataset in: *6. Maps\6.2. map for 2023\1. data\LAU_RG_01M_2023_4326.gpkg*

The GeoPackage is the Eurostat LAU (Local Administrative Units) 2023 layer. It
covers every country in the dataset — the `geom` column already comes in EPSG 4326.

```{r}
lau_2023 <- st_read(
  here('6. Maps', '6.2. map for 2023', '1. data', 'LAU_RG_01M_2023_4326.gpkg'),
  quiet = TRUE
)

glimpse(lau_2023)
```



# 2. create the dataset
> Here I want to create the dataset based on the instrcutions at the beginning of the document in *What I want*. Save the new file in: *6. Maps\6.2. map for 2023\2. final dataset*

Keep only Germany (`CNTR_CODE == 'DE'`). The 8-digit German municipality code
(AGS) is derived from `GISCO_ID` by dropping the `DE_` prefix — same rule used
in *2. generate the maps.qmd*. Population, density and area are kept as the
"additional variables" that may be dropped later.

Create the dataset
```{r}
municipalities_2023 <- lau_2023 |>
  filter(CNTR_CODE == 'DE') |>
  mutate(MunicipalityCode = sub('DE_', '', GISCO_ID)) |>
  select(
    Municipality     = LAU_NAME,
    MunicipalityCode,
    Population2023   = POP_2023,
    PopDensity2023   = POP_DENS_2023,
    AreaKm2          = AREA_KM2,
    geom
  )

municipalities_2023
```

Save the dataset as a (geo)parquet file, matching the format of the other
spatial datasets in *6. Maps\6.1. Dataset*.

```{r}
st_write_parquet(
  municipalities_2023,
  here('6. Maps', '6.2. map for 2023', '2. final dataset', 'municipalities.parquet')
)
```




# 3. check the dataset
> Here I want to perform basic checks and create a simple map (using leaflet function) to show the municipalities all across Germany. Additionaly, you can include the bounderies of the Federal states (here you have the file: *6. Maps\6.1. Dataset\federalState.parquet*). In order to do that, you can simulate a variable and display it.

## 3.1. Basic checks

Row count, missing values, code uniqueness, CRS and geometry validity.

```{r}
data.frame(
  Rows           = nrow(municipalities_2023),
  MissingName    = sum(is.na(municipalities_2023$Municipality)),
  MissingCode    = sum(is.na(municipalities_2023$MunicipalityCode)),
  UniqueCodes    = dplyr::n_distinct(municipalities_2023$MunicipalityCode),
  CodeWidth8     = all(nchar(municipalities_2023$MunicipalityCode) == 8),
  CRS            = st_crs(municipalities_2023)$input,
  GeomType       = paste(unique(as.character(st_geometry_type(municipalities_2023))), collapse = ', '),
  AllValid       = all(st_is_valid(municipalities_2023))
)

glimpse(municipalities_2023)
```

## 3.2. Map

Load the federal-state boundaries and simulate a numeric variable so the
municipalities can be coloured.

```{r}
federal_states <- st_read_parquet(
  here('6. Maps', '6.1. Dataset', 'federalState.parquet')
) |>
  st_transform(4326)

set.seed(123)
municipalities_map <- municipalities_2023 |>
  st_transform(4326) |>
  mutate(sim_value = runif(n()))
```

Leaflet map — municipalities filled by the simulated variable, federal-state
borders drawn on top.

```{r}
pal <- colorNumeric('YlOrRd', domain = municipalities_map$sim_value)

leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    data        = municipalities_map,
    fillColor   = ~pal(sim_value),
    weight      = 0.2,
    color       = 'white',
    fillOpacity = 0.7,
    label       = ~Municipality,
    popup       = ~paste0('<strong>', Municipality, '</strong><br>',
                          'Code: ', MunicipalityCode, '<br>',
                          'Simulated value: ', round(sim_value, 3))
  ) |>
  addPolygons(
    data        = federal_states,
    fill        = FALSE,
    weight      = 1.5,
    color       = 'black',
    label       = ~FederalState
  ) |>
  addLegend(
    data     = municipalities_map,
    pal      = pal,
    values   = ~sim_value,
    title    = 'Simulated value',
    position = 'bottomright'
  )
```



# 4. Population data
> Here I want to create and load population data for all the municiplaities in Germany in 2023 and find how many I can find a match.


Build the dataset for the municiplaity population data:

```{r}
#Credentials:
# gen_auth_save(database = 'genesis')
# gen_auth_save(database = 'regio')

gen_logincheck(database = "genesis")
gen_logincheck(database = "regio")


# Load the data:
pop.muni.2023 <- gen_table(
  name      = "12411-01-01-5",
  database  = "regio",
  language  = "en",
  startyear = 2023,
  endyear   = 2023
) |>
  filter(`1_variable_code` == "GEMEIN",          # municipalities only (drops DINSG/DLAND/REGBEZ/KREISE)
         is.na(`2_variable_attribute_code`)) |>  # sex total
  select(
    MunicipalityCode = `1_variable_attribute_code`,
    Municipality     = `1_variable_attribute_label`,
    year             = time,
    population       = value
  ) |>
  mutate(
    date       = as.Date(paste0(substr(year, 1, 4), "-12-31")),
    population = suppressWarnings(as.numeric(population))
  ) |>
  select(Municipality, MunicipalityCode, date, population) 
```

Here I want to see if there is a match:

```{r}
# Join the two population sources on the 8-digit municipality code (AGS):
#   municipalities_2023 -> POP_2023 from the Eurostat LAU layer
#   pop.muni.2023       -> population from Destatis regio table 12411-01-01-5
pop_compare <- municipalities_2023 |>
  st_drop_geometry() |>
  select(MunicipalityCode,
         Municipality_LAU = Municipality,
         Population_LAU   = Population2023) |>
  full_join(
    pop.muni.2023 |>
      select(MunicipalityCode,
             Municipality_API = Municipality,
             Population_API   = population),
    by = "MunicipalityCode"
  ) |>
  mutate(
    in_LAU       = !is.na(Municipality_LAU),
    in_API       = !is.na(Municipality_API),
    matched      = in_LAU & in_API,
    pop_diff     = Population_API - Population_LAU,
    pop_pct_diff = round(100 * pop_diff / Population_LAU, 1)
  )

# Show the largest population discrepancies among the matched municipalities
pop_compare |>
  arrange(desc(abs(pop_pct_diff)))
```


Here I want to know how many matches:

```{r}
pop_compare |>
  summarise(
    LAU_municipalities = sum(in_LAU),                 # rows from the LAU geometry
    API_municipalities = sum(in_API),                 # rows from the Destatis API
    matched            = sum(matched),                # code present in both
    only_in_LAU        = sum(in_LAU & !in_API),       # geometry, no population
    only_in_API        = sum(in_API & !in_LAU),       # population, no geometry
    match_rate_LAU     = paste0(round(100 * sum(matched) / sum(in_LAU), 1), "%"),
    # population agreement on the matched municipalities
    identical_pop      = sum(pop_diff[matched] == 0, na.rm = TRUE),
    mean_abs_pct_diff  = round(mean(abs(pop_pct_diff[matched]), na.rm = TRUE), 1)
  )
```