---
title: 'Federal-state PV registration dataset'
format: html
---

> This notebook builds the daily dataset used by the federal-state Shiny app. For each of Germany's 16 federal states and each day, it counts new PV-unit registrations from the German MaStR registry, splits them into balcony and non-balcony systems, and joins federal-state population so counts can also be expressed per 100,000 inhabitants.

**Inputs**:

- MaStR snapshot — `EinheitenSolar.parquet` (German solar-unit registry).
- District-code lookup — `districts.parquet`.
- Federal-state population — Destatis GENESIS table `12411-0010`, fetched via the `restatis` package.

**Output** (saved next to the Shiny app):

- `state_day.parquet` — daily registrations per federal state × PV category, with population and per-100k variants.

**Libraries**:

```{r}
pacman::p_load(
  here,        # here()
  arrow,       # open_dataset(), read_parquet(), write_parquet(), collect()
  sfarrow,     # st_read_parquet()
  sf,          # st_drop_geometry()
  dplyr,       # rename, mutate, case_when, if_else, left_join, relocate,
               # filter, select, group_by, summarise, arrange, pull, ungroup, n
  tibble,      # tibble()
  tidyr,       # complete()
  lubridate,   # floor_date()
  stringr,     # str_pad()
  restatis,    # gen_logincheck(), gen_table()
  shiny        # shinyApp(), fluidPage(), selectizeInput(), updateSelectizeInput(), renderPrint(), verbatimTextOutput(), req()
)
```


**Frequency**: daily registrations.

# 1. Build the dataset

**Step 1 — Load and clean the MaStR snapshot** (rename German columns, recoerce types, attach lookup tables, derive the balcony/non-balcony category, filter to Germany, add date helpers):


```{r}
data <- open_dataset(here("01_data/1.2. May 7 2026/EinheitenSolar.parquet"))


# Step 1: rename German → English (explicit mapping, column order differs from old dataset)
data <- data |>
  rename(
    UnitMastrNumber                          = EinheitMastrNummer,
    DateLastUpdate                           = DatumLetzteAktualisierung,
    LocationMastrNumber                      = LokationMaStRNummer,
    GridOperatorVerificationStatus           = NetzbetreiberpruefungStatus,
    GridOperatorVerificationDate             = NetzbetreiberpruefungDatum,
    PlantOperatorMastrNumber                 = AnlagenbetreiberMastrNummer,
    Country                                  = Land,
    FederalState                             = Bundesland,
    District                                 = Landkreis,
    Municipality                             = Gemeinde,
    MunicipalityCode                         = Gemeindeschluessel,
    PostalCode                               = Postleitzahl,
    City                                     = Ort,
    RegistrationDate                         = Registrierungsdatum,
    CommissioningDate                        = Inbetriebnahmedatum,
    UnitSystemStatus                         = EinheitSystemstatus,
    UnitOperationalStatus                    = EinheitBetriebsstatus,
    NotPresentInMigratedUnits                = NichtVorhandenInMigriertenEinheiten,
    PowerGenerationUnitName                  = NameStromerzeugungseinheit,
    Weic_NotAvailable                        = Weic_nv,
    PowerPlantNumber_NotAvailable            = Kraftwerksnummer_nv,
    EnergySource                             = Energietraeger,
    GrossCapacity                            = Bruttoleistung,
    NetNominalCapacity                       = Nettonennleistung,
    RemoteControllabilityGridOperator        = FernsteuerbarkeitNb,
    FeedInType                               = Einspeisungsart,
    AssignedActivePowerInverter              = ZugeordneteWirkleistungWechselrichter,
    NumberOfModules                          = AnzahlModule,
    TypeOfSolarSystem                        = ArtDerSolaranlage,
    PowerLimitation                          = Leistungsbegrenzung,
    UniformOrientationAndTiltAngle           = EinheitlicheAusrichtungUndNeigungswinkel,
    MainOrientation                          = Hauptausrichtung,
    MainOrientationTiltAngle                 = HauptausrichtungNeigungswinkel,
    UsageArea                                = Nutzungsbereich,
    EegMastrNumber                           = EegMaStRNummer,
    StorageAtSameLocation                    = SpeicherAmGleichenOrt,
    SecondaryOrientation                     = Nebenausrichtung,
    SecondaryOrientationTiltAngle            = NebenausrichtungNeigungswinkel,
    Street                                   = Strasse,
    StreetNotFound                           = StrasseNichtGefunden,
    HouseNumber_NotAvailable                 = Hausnummer_nv,
    HouseNumberNotFound                      = HausnummerNichtGefunden,
    Longitude                                = Laengengrad,
    Latitude                                 = Breitengrad,
    DateStartTemporaryDecommissioning        = DatumBeginnVoruebergehendeStilllegung,
    HouseNumber                              = Hausnummer,
    AddressAddition                          = Adresszusatz,
    RemoteControllabilityDirectMarketer      = FernsteuerbarkeitDv,
    PreviousPlantOperatorMastrNumber         = AltAnlagenbetreiberMastrNummer,  # new variable
    DateOfOperatorChange                     = DatumDesBetreiberwechsels,
    DateRegistrationOfOperatorChange         = DatumRegistrierungDesBetreiberwechsels,
    CadastralDistrict                        = Gemarkung,
    FieldParcelNumbers                       = FlurFlurstuecknummern,
    ConnectionToExtraHighOrHighVoltage       = AnschlussAnHoechstOderHochSpannung,
    DatePermanentDecommissioning             = DatumEndgueltigeStilllegung,
    SizeOfUsedAreaInHectares                 = GroesseDerInAnspruchGenommenenFlaecheInHektar,
    PreviousUsageTypeOfArea                  = VorherigerNutzungsartenbereichDerFlaeche,
    PredominantUsageTypeBeforeSolarConstruction = UeberwiegendeNutzungsartDerFlaecheVorErrichtung,
    NameOfSolarPark                          = NameDesSolarparks,
    AdditionalFeaturesOfAreaAndCurrentUsage  = ZusaetzlicheMerkmaleDerFlaecheUndDerAktuellenFlaechennutzung,
    GenMastrNumber                           = GenMastrNummer,
    CommissioningDateAtCurrentLocation       = InbetriebnahmedatumAmAktuellenStandort,
    DispatchResponsible                      = Einsatzverantwortlicher,
    CitizenEnergy                            = Buergerenergie,
    DateResumptionOfOperation                = DatumWiederaufnahmeBetrieb,
    PlannedCommissioningDate                 = GeplantesInbetriebnahmedatum,
    ClearanceHeight                          = LichteHoehe,
    PowerPlantNumber                         = Kraftwerksnummer
  )

# Step 2: convert numeric columns (XML parsing stores everything as character)
data <- data |>
  mutate(
    TypeOfSolarSystem           = as.integer(TypeOfSolarSystem),
    Country                     = as.integer(Country),
    FederalState                = as.integer(FederalState),
    GrossCapacity               = as.numeric(GrossCapacity),
    NetNominalCapacity          = as.numeric(NetNominalCapacity),
    AssignedActivePowerInverter = as.numeric(AssignedActivePowerInverter),
    NumberOfModules             = as.integer(NumberOfModules)
  )

# Step 3: lookup tables
C1 <- data.frame(
  TypeOfSolarSystem.names = c(
    "Structural_facilities", "Other_Structural_facilities", "Open_space",
    "Bodies_water", "Large_parking_lot", "Plug_in_sola_system", "Onshore&Offshore_wind_power"
  ),
  TypeOfSolarSystem = as.integer(c(853, 2484, 852, 3002, 3058, 2961, NA))
)

C2 <- tibble(
  Country = as.integer(c(84, 231, 90, 66, 198, 206, 106)),
  Country.sn = c("Germany", "Switzerland", "Denmark", "Belgium", "Netherlands", "Austria", "France")
)

C3 <- data.frame(
  FederalState = as.integer(c(
    1409, 1402, 1400, 1403, 1412, 1408, 1410, 1407,
    1414, 1411, 1413, 1405, 1415, 1406, 1401, 1404
  )),
  FederalState.names = c(
    "North Rhine-Westphalia", "Baden-Württemberg", "Brandenburg", "Bavaria",
    "Saarland", "Lower Saxony", "Rhineland-Palatinate", "Mecklenburg-Western Pomerania",
    "Saxony-Anhalt", "Schleswig-Holstein", "Saxony", "Hesse",
    "Thuringia", "Hamburg", "Berlin", "Bremen"
  ),
  FederalStateCode = c(
    "DEA", "DE1", "DE4", "DE2",
    "DEC", "DE9", "DEB", "DEM",
    "DEE", "DEF", "DED", "DE7",
    "DEG", "DE6", "DE3", "DE5"
  )
)

# district code from map (no geometry needed)
districts_map <- st_read_parquet(here("ignore/3. data/2. maps/districts.parquet")) |>
  st_drop_geometry() |>
  select(District, DistrictCode)

# Step 4: join lookups
data <- data |>
  left_join(C1, by = "TypeOfSolarSystem") |>
  relocate(TypeOfSolarSystem.names, .after = TypeOfSolarSystem) |>
  left_join(C2, by = "Country") |>
  relocate(Country.sn, .after = Country) |>
  left_join(C3, by = "FederalState") |>
  relocate(FederalState.names, .after = FederalState) |>
  relocate(FederalStateCode, .after = FederalState.names) |>
  left_join(districts_map, by = "District") |>
  relocate(DistrictCode, .after = District)

# Step 5: solar_category
reform_date <- as.Date("2024-05-16")

data <- data |>
  mutate(
    RegistrationDate  = as.Date(RegistrationDate),
    CommissioningDate = as.Date(CommissioningDate),
    solar_category = case_when(
      TypeOfSolarSystem == 2961 & RegistrationDate <  reform_date & AssignedActivePowerInverter <= 0.6 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 & RegistrationDate >= reform_date & AssignedActivePowerInverter <= 0.8 & GrossCapacity <= 2.0 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 ~ "Plug_in_sola_system - Non-Balcony",
      TRUE ~ TypeOfSolarSystem.names
    )
  )

# Step 6: filter Germany
data <- data |>
  filter(Country.sn == "Germany")

# Step 7: date variables (week, month, quarter, year)
data <- data |>
  mutate(
    date.registration.day     = RegistrationDate,
    date.registration.week    = floor_date(RegistrationDate,  "week"),
    date.registration.month   = floor_date(RegistrationDate,  "month"),
    date.registration.quarter = floor_date(RegistrationDate,  "quarter"),
    date.registration.year    = floor_date(RegistrationDate,  "year"),
    date.commission.day       = CommissioningDate,
    date.commission.week      = floor_date(CommissioningDate, "week"),
    date.commission.month     = floor_date(CommissioningDate, "month"),
    date.commission.quarter   = floor_date(CommissioningDate, "quarter"),
    date.commission.year      = floor_date(CommissioningDate, "year")
  )

# Step 8: fix MunicipalityCode
data <- data |>
  mutate(MunicipalityCode = str_pad(as.character(MunicipalityCode), 8, pad = "0"))

# Collapse every category other than `Balcony PVs` into a single `Non-balcony PVs` group:


data <- data |>
  mutate(solar_category = if_else(solar_category == "Plug_in_sola_system - Balcony",
                                  "Balcony PVs",
                                  "Non-balcony PVs"))


```

**Step 2 — Aggregate to daily counts per federal state and PV category** (fill missing date slots with zeros and compute cumulative totals):

```{r}
state_day <- data |>
  group_by(FederalStateCode, FederalState.names, solar_category, date.registration.day) |>
  summarise(n = n(), .groups = "drop") |>
  collect() |>
  complete(nesting(FederalStateCode, FederalState.names), solar_category,
           date.registration.day = seq(min(date.registration.day), max(date.registration.day), by = "day"),
           fill = list(n = 0)) |>
  arrange(FederalStateCode, solar_category, date.registration.day) |>
  group_by(FederalStateCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```



**Step 3 — Federal-state population** (from Destatis GENESIS table `12411-0010`, reference date 31 December). A `restatis` GENESIS account is required; the `gen_auth_save()` calls below need to be run once to store credentials:

```{r}
# Credentials
# gen_auth_save(database = "genesis")   # for www-genesis.destatis.de
# gen_auth_save(database = "regio")     # for www.regionalstatistik.de
# gen_auth_save(database = "zensus")    # for ergebnisse.zensus2022.de (if you ever need it)

gen_logincheck(database = "genesis")
gen_logincheck(database = "regio")

# Create the dataset
pop.federal_state <- gen_table(
  name      = "12411-0010",
  database  = "genesis",
  language  = "en",
  startyear = 2023,
  endyear   = 2023
)

federal_state.codes <- tibble::tibble(
  code = c("01","02","03","04","05","06","07","08","09","10","11","12","13","14","15","16"),
  nuts = c("DEF","DE6","DE9","DE5","DEA","DE7","DEB","DE1","DE2","DEC","DE3","DE4","DEM","DED","DEE","DEG")
)

pop.federal_state.2023 <- pop.federal_state |>
  select(
    code       = `1_variable_attribute_code`,
    name       = `1_variable_attribute_label`,
    year       = time,
    population = value
  ) |>
  mutate(
    date       = as.Date(paste0(substr(year, 1, 4), "-12-31")),
    population = suppressWarnings(as.numeric(population))
  ) |>
  select(-year) |>
  left_join(federal_state.codes, by = "code") |>
  select(FederalState.names = name, FederalStateCode = nuts, date, population) |>
  arrange(FederalStateCode, date)
```


**Step 4 — Merge population into the daily series** (creates per-100k variants):
```{r}
pop_lookup <- pop.federal_state.2023 |>
  select(FederalStateCode, population)

state_day <- state_day |>
  left_join(pop_lookup, by = "FederalStateCode") |>
  mutate(
    n_per100k            = (n            / population) * 100000,
    n_cumulative_per100k = (n_cumulative / population) * 100000
  )
```

**Step 5 — Save** the daily dataset next to the Shiny app:

```{r}
out_dir <- here("8. shiny app", "8.1. shiny apps", "4")

write_parquet(state_day, file.path(out_dir, "state_day.parquet"))
```

**Load** (use this to reopen the saved dataset later without re-running the pipeline):

```{r}
out_dir <- here("8. shiny app", "8.1. shiny apps", "4")

state_day <- read_parquet(file.path(out_dir, "state_day.parquet"))
```

**Sanity check** — daily registrations per state, faceted by PV category:

```{r}
state_day |>
  ggplot(aes(x = date.registration.day, y = n, color = FederalState.names)) +
  geom_line() +
  facet_wrap(~solar_category)
```