---
title: 'Shiny app for municipalities'
format: html
---



::: {.callout}
**Shiny app I want to create**:
- I want to create a shiny app where on the left there is a leaflet map of all federla states bounderies and the outcome variable can be (there are different options, and you can choose which one you want):
  1. cumulative registrations from the beginning of the sample to the end (not just 2026, but also the other years). You can use either population adjusted or not
  2. (perhaps) the growth rate of registrations from before Iran war shock and after. You can decide how to calculate this growth rate, but it should be something that can capture the change in the adoption rate of solar units after the shock of the Iran war. 
  3. suggest me any relevant measure I can use as outcome variable to capture the shock of the Iran war on the adoption rate of solar units at the federal state level. I want to have the option to choose which one I want to use as outcome variable in the map

Idea of the map: identify which federal state (by looking at the map) was hit more by the Iran war or the shock of the iran war can have a larger effect.


Also, you have to add an option allowing you to decide whether you want population adjusted or not


- on the right top there must a chart of the daily registration in grey + MA
Once you seelct on the map one federal state, it dispays the chart of that specific state of the daily registrations (popuation corrected) on the backgroung in grey + moving average (with a slide to deicde the window)
As default federal state to show at the beginning, even before seelcting a specific state, show the plot for Baden wruttemberg federal state
in the plot add the vertcial line sin black for the iran war relevant dates (max 2 dates, no more. And they must be relevant for the impact of the Iran war on the adoption rate of solar units)
I want to have a slider (only for this chart) where you can select the range of the period to choose.
Keep the scale of y-axis the same across federal states (so that once i click on a different federal state it shows that with the same scale of the other one)

- on the right down there must be a plot of the selected federal state that show the MA of daily registration for each year (so that you have for each year the MA of daily registration). Be careful that for 2026 there is no data after May. Decide if for the other years I should dispaly the whole series or I should cap it up to May.
There must be also a slider in the shiny app that you can use to chnage the window
I want to show a red color for 2026 and a different group of colors for all the others (i must be in any case be able to distinguish between the other years)
Keep the scale of y-axis the same across federal states (so that once i click on a different federal state it shows that with the same scale of the other one)

- create an option where you can decide whether to focs on balcoy PVs or non-balcoy PVs

Here you have the datasets:
  1. dataset for federal states bounderies: *8. shiny app\8.1. shiny apps\4\federalState.parquet*
  3. the dataset where you cna find the registration is here: *8. shiny app\8.1. shiny apps\4\state_day.parquet*

Here (*8. shiny app\8.1. shiny apps\4\1. dataset.qmd*) you can find the file i used to create the dataset for the balcony PVs registrations
I want you to create a shiny app, where you include both the user interface and the server in the same chunk.

:::




**Libraries**:

```{r}
pacman::p_load(
  shiny, leaflet, sf, sfarrow, arrow,
  dplyr, tidyr, ggplot2, here, tibble, lubridate,
  bslib, htmltools
)
```

Shiny app:

```{r}
# ============================================================================
# Federal-state PV adoption in Germany — cumulative-registration map + charts
# ============================================================================

# ---- 0. data ---------------------------------------------------------------
base_dir <- here("8. shiny app", "8.1. shiny apps", "4")

fs_sf <- st_read_parquet(file.path(base_dir, "federalState.parquet")) |>
  rename(FederalState.names = FederalState) |>
  st_zm(drop = TRUE) |>
  st_transform(4326) |>
  st_make_valid()
fs_sf <- fs_sf[!st_is_empty(fs_sf), ]

state_day <- read_parquet(file.path(base_dir, "state_day.parquet")) |>
  mutate(date.registration.day = as.Date(date.registration.day),
         year = year(date.registration.day))

# ---- 1. constants ----------------------------------------------------------
default_state <- "DE1"                       # Baden-Württemberg

# Only 2 dates — war begins (treatment onset) and ceasefire (treatment ends).
# These bracket the Iran-war shock window for PV-adoption analysis.
iran_milestones <- tribble(
  ~date,                  ~label,
  as.Date("2026-02-28"),  "Iran war begins",
  as.Date("2026-04-08"),  "Ceasefire"
)

# Cap all years at May 7 — 2026 has no data past that, so the per-year
# comparison is apples-to-apples only up to that cut-off.
may7 <- as.Date("2000-05-07")

# date bounds for the daily-chart range slider
date_min <- min(state_day$date.registration.day, na.rm = TRUE)
date_max <- max(state_day$date.registration.day, na.rm = TRUE)

# Iran-war shock parameters used by the map-metric ratios
iran_war_start <- as.Date("2026-02-28")
shock_window   <- 30L                            # days on each side

# ---- 2. UI -----------------------------------------------------------------
ui <- page_fluid(
  theme = bs_theme(
    version    = 5,
    bootswatch = "flatly",
    primary    = "#2d4a7e"
  ),
  tags$head(tags$style(HTML("
    body { background: #f5f6f8; }
    .card { border: none; border-radius: 0.6rem;
            box-shadow: 0 2px 8px rgba(0,0,0,0.05); margin-bottom: 16px; }
    .card-header { background: #ffffff; border-bottom: 1px solid #e9ecef;
                   font-weight: 600; padding: 0.8rem 1.1rem; }
    .leaflet-container { border-radius: 0 0 0.6rem 0.6rem; }
    .irs--shiny .irs-bar { background: #2d4a7e; border-color: #2d4a7e; }
    .irs--shiny .irs-handle { border-color: #2d4a7e; }
    .irs--shiny .irs-from, .irs--shiny .irs-to, .irs--shiny .irs-single {
      background: #2d4a7e;
    }
    .form-check-input:checked { background-color: #2d4a7e;
                                border-color: #2d4a7e; }
  "))),

  # gradient title bar — colours match the map's blue → orange ramp
  div(
    style = paste0(
      "background: linear-gradient(90deg, #2d4a7e 0%, #b65523 100%); ",
      "color: white; padding: 22px 28px; border-radius: 0.6rem; ",
      "margin-bottom: 18px; box-shadow: 0 2px 8px rgba(0,0,0,0.05);"
    ),
    h2("Iran war & federal-state PV adoption — Germany",
       style = "margin: 0; font-weight: 700; font-size: 1.6rem;"),
    p(paste0("Daily PV-unit registrations from the MaStR registry, by ",
             "German federal state. Click a state on the map to inspect ",
             "its time series. Y-axis scales are held fixed across states ",
             "for the current PV-type / outcome choice."),
      style = "margin: 6px 0 0 0; opacity: 0.92; font-size: 0.95rem;")
  ),

  # control panel
  card(
    card_body(
      fluidRow(
        column(3, radioButtons("category", "PV type:",
                               choices  = c("Balcony PVs", "Non-balcony PVs"),
                               selected = "Balcony PVs")),
        column(3, radioButtons("adjust", "Outcome:",
                               choices  = c("Per 100k inhab." = "pop",
                                            "Raw count"       = "raw"),
                               selected = "pop")),
        column(3, sliderInput("ma_window_daily",
                              "MA window — daily chart (days):",
                              min = 3, max = 60, value = 14, step = 1)),
        column(3, sliderInput("ma_window_yearly",
                              "MA window — yearly chart (days):",
                              min = 3, max = 60, value = 14, step = 1))
      ),
      fluidRow(
        column(6, radioButtons(
          "map_metric", "Map outcome:",
          choices = c(
            "Cumulative registrations"                  = "cumulative",
            "Post/Pre ratio (±30d around war start)"    = "prepost",
            "YoY same-window ratio (2026 / 2025)"       = "yoy"
          ),
          selected = "cumulative",
          inline   = TRUE
        )),
        column(6, sliderInput("daily_range", "Date range — daily chart:",
                              min   = date_min,
                              max   = date_max,
                              value = c(date_min, date_max),
                              timeFormat = "%Y-%m-%d",
                              width = "100%"))
      )
    )
  ),

  # map + charts
  layout_columns(
    col_widths = c(6, 6),
    card(
      card_header("Map of federal states"),
      full_screen = TRUE,
      leafletOutput("map", height = 700)
    ),
    card(
      card_header("Selected federal-state time series"),
      full_screen = TRUE,
      plotOutput("daily_chart",  height = "345px"),
      plotOutput("yearly_chart", height = "345px")
    )
  )
)

# ---- 3. server -------------------------------------------------------------
server <- function(input, output, session) {

  selected_state <- reactiveVal(default_state)

  # data for the chosen PV category
  cat_data <- reactive({
    state_day |> filter(solar_category == input$category)
  })

  # per-federal-state map metric — one of three outcomes (see UI selector).
  # `value` is the column that drives the choropleth fill.
  metric_tbl <- reactive({
    cdf <- cat_data()
    m   <- input$map_metric

    if (m == "cumulative") {
      val_col <- if (input$adjust == "pop") "n_per100k" else "n"
      cdf |>
        group_by(FederalStateCode) |>
        summarise(value = sum(.data[[val_col]], na.rm = TRUE),
                  .groups = "drop")

    } else if (m == "prepost") {
      # mean daily registrations in the 30 days AFTER the war start, divided
      # by the mean in the 30 days BEFORE. Population cancels in the ratio,
      # so the per-100k vs raw toggle is irrelevant here.
      pre_lo  <- iran_war_start - shock_window
      pre_hi  <- iran_war_start - 1
      post_lo <- iran_war_start
      post_hi <- iran_war_start + shock_window - 1
      cdf |>
        group_by(FederalStateCode) |>
        summarise(
          pre  = mean(n[date.registration.day >= pre_lo  &
                        date.registration.day <= pre_hi],  na.rm = TRUE),
          post = mean(n[date.registration.day >= post_lo &
                        date.registration.day <= post_hi], na.rm = TRUE),
          value = if_else(pre > 0, post / pre, NA_real_),
          .groups = "drop"
        ) |>
        select(FederalStateCode, value)

    } else {  # "yoy"
      # 30-day post-shock window in 2026 vs the same calendar window in 2025.
      # Controls for the seasonal pattern of PV registrations.
      win_2025_lo <- iran_war_start - years(1)
      win_2025_hi <- iran_war_start - years(1) + shock_window - 1
      win_2026_lo <- iran_war_start
      win_2026_hi <- iran_war_start + shock_window - 1
      cdf |>
        group_by(FederalStateCode) |>
        summarise(
          y2025 = sum(n[date.registration.day >= win_2025_lo &
                        date.registration.day <= win_2025_hi], na.rm = TRUE),
          y2026 = sum(n[date.registration.day >= win_2026_lo &
                        date.registration.day <= win_2026_hi], na.rm = TRUE),
          value = if_else(y2025 > 0, y2026 / y2025, NA_real_),
          .groups = "drop"
        ) |>
        select(FederalStateCode, value)
    }
  })

  map_data <- reactive({
    fs_sf |> left_join(metric_tbl(), by = "FederalStateCode")
  })

  # global y-max for the daily chart — max across all states for the current
  # category + outcome, so the y-axis is fixed when the user clicks a different
  # state. Recomputed when category or outcome changes.
  # Hard cap at 6 for Balcony PVs + per-100k so the daily-bar spikes don't
  # dominate the moving-average line.
  daily_y_max <- reactive({
    val_col <- if (input$adjust == "pop") "n_per100k" else "n"
    if (input$category == "Balcony PVs" && input$adjust == "pop") {
      6
    } else {
      suppressWarnings(max(cat_data()[[val_col]], na.rm = TRUE))
    }
  })

  # global y-max for the yearly chart — max of the per-year MA across all
  # states, for the current category + outcome + yearly MA window.
  yearly_y_max <- reactive({
    val_col <- if (input$adjust == "pop") "n_per100k" else "n"
    k <- input$ma_window_yearly
    cat_data() |>
      mutate(md = make_date(2000,
                            month(date.registration.day),
                            day(date.registration.day))) |>
      filter(md <= may7) |>
      arrange(FederalStateCode, year, md) |>
      group_by(FederalStateCode, year) |>
      mutate(ma = as.numeric(stats::filter(.data[[val_col]],
                                           rep(1 / k, k), sides = 1))) |>
      ungroup() |>
      pull(ma) |>
      max(na.rm = TRUE)
  })

  # base map: tiles + panes (rendered once)
  output$map <- renderLeaflet({
    leaflet() |>
      addProviderTiles(providers$CartoDB.Positron) |>
      setView(lng = 10.45, lat = 51.16, zoom = 6) |>
      addMapPane("choropleth", zIndex = 410) |>
      addMapPane("highlight",  zIndex = 420)
  })

  # choropleth — redrawn when category, outcome, or map metric changes
  observe({
    md  <- map_data()
    pal <- colorNumeric(
      palette  = c("#2d4a7e", "#4e74a3", "#92acc7", "#d4cbb5",
                   "#e2c79a", "#e3a978", "#d7864c", "#b65523"),
      domain   = md$value,
      na.color = "#bbbbbb"
    )

    # legend title and hover-label format depend on the chosen metric
    legend_title <- switch(input$map_metric,
      "cumulative" = if (input$adjust == "pop") "Cumulative<br>per 100k"
                     else                       "Cumulative<br>count",
      "prepost"    = "Post / Pre ratio<br>(±30 d window)",
      "yoy"        = "2026 / 2025<br>(30 d post-shock)"
    )
    md$label_text <- if (input$map_metric == "cumulative") {
      ifelse(is.na(md$value),
             paste0(md$FederalState.names, ": no data"),
             paste0(md$FederalState.names, ": ",
                    format(round(md$value), big.mark = ",")))
    } else {
      ifelse(is.na(md$value),
             paste0(md$FederalState.names, ": no data"),
             paste0(md$FederalState.names, ": ",
                    sprintf("%.2fx", md$value)))
    }

    leafletProxy("map") |>
      clearGroup("state") |>
      clearControls() |>
      addPolygons(
        data        = md, group = "state",
        layerId     = ~FederalStateCode,
        fillColor   = ~pal(value),
        fillOpacity = 0.85,
        color       = "#ffffff",
        weight      = 0.6,
        options     = pathOptions(pane = "choropleth"),
        label       = ~label_text,
        highlightOptions = highlightOptions(weight = 1.5, color = "#000000",
                                            bringToFront = FALSE)
      ) |>
      addLegend(position = "bottomright", pal = pal, values = md$value,
                title = HTML(legend_title))
  })

  # click → set selection
  observeEvent(input$map_shape_click, {
    id <- input$map_shape_click$id
    if (!is.null(id)) selected_state(id)
  })

  # highlight selected state
  observe({
    sel <- fs_sf |> filter(FederalStateCode == selected_state())
    p   <- leafletProxy("map") |> clearGroup("sel")
    if (nrow(sel) == 1) {
      p |> addPolygons(data = sel, group = "sel", fill = FALSE,
                       color = "#1f4e79", weight = 3,
                       options = pathOptions(pane = "highlight"))
    }
  })

  # ---- daily chart (full sample) -------------------------------------------
  output$daily_chart <- renderPlot({
    df <- cat_data() |>
      filter(FederalStateCode == selected_state(),
             date.registration.day >= input$daily_range[1],
             date.registration.day <= input$daily_range[2]) |>
      arrange(date.registration.day)
    validate(need(nrow(df) > 0, "No data for this state in the selected range."))

    sname <- df$FederalState.names[1]
    df$y  <- if (input$adjust == "pop") df$n_per100k else df$n

    k     <- input$ma_window_daily
    df$ma <- as.numeric(stats::filter(df$y, rep(1 / k, k), sides = 1))

    ms <- iran_milestones |>
      filter(date >= input$daily_range[1],
             date <= input$daily_range[2])
    y_max <- daily_y_max()
    y_lab <- if (input$adjust == "pop") "Daily reg. per 100k"
             else                       "Daily reg. (raw count)"

    ggplot(df, aes(x = date.registration.day)) +
      geom_col(aes(y = y), fill = "grey70", width = 1) +
      geom_vline(data = ms, aes(xintercept = date),
                 color = "black", linetype = "dashed", linewidth = 0.5) +
      geom_text(data = ms, aes(x = date, y = y_max, label = label),
                angle = 90, hjust = 1, vjust = -0.3, size = 3,
                color = "grey20", check_overlap = TRUE) +
      geom_line(aes(y = ma), color = "#c0392b", linewidth = 0.9, na.rm = TRUE) +
      scale_x_date(date_labels = "%Y-%m", date_breaks = "6 months") +
      coord_cartesian(ylim = c(0, y_max * 1.05)) +
      labs(title    = paste0(sname, " — daily PV registrations"),
           subtitle = paste(input$category, "•", k, "-day MA"),
           x = NULL, y = y_lab) +
      theme_minimal(base_size = 12) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  })

  # ---- per-year chart (MA, Jan–May 7) --------------------------------------
  output$yearly_chart <- renderPlot({
    df <- cat_data() |>
      filter(FederalStateCode == selected_state()) |>
      mutate(md = make_date(2000,
                            month(date.registration.day),
                            day(date.registration.day))) |>
      filter(md <= may7) |>
      arrange(year, md)
    validate(need(nrow(df) > 0, "No data for this state."))

    sname <- df$FederalState.names[1]
    df$y  <- if (input$adjust == "pop") df$n_per100k else df$n
    k     <- input$ma_window_yearly
    df <- df |>
      group_by(year) |>
      mutate(ma = as.numeric(stats::filter(y, rep(1 / k, k), sides = 1))) |>
      ungroup()

    # year palette: 2026 = red, other years = viridis (still distinguishable).
    years_chr <- sort(unique(as.character(df$year)))
    other_yrs <- setdiff(years_chr, "2026")
    year_pal  <- setNames(
      hcl.colors(max(length(other_yrs), 1), palette = "Viridis", rev = FALSE),
      other_yrs
    )
    if ("2026" %in% years_chr) year_pal["2026"] <- "red"

    # 2026 last in plotting order so it sits on top of the other lines
    df$year_f <- factor(df$year, levels = c(other_yrs, "2026"))

    y_lab <- if (input$adjust == "pop") "MA daily reg. per 100k"
             else                       "MA daily reg."

    ggplot(df, aes(x = md, y = ma, color = year_f, group = year_f)) +
      geom_line(linewidth = 0.8, na.rm = TRUE) +
      scale_x_date(date_labels = "%b", date_breaks = "1 month") +
      scale_color_manual(values = year_pal, name = "Year",
                         breaks = years_chr) +
      coord_cartesian(ylim = c(0, yearly_y_max() * 1.05)) +
      labs(title    = paste0(sname, " — MA daily registrations by year (Jan–May 7)"),
           subtitle = paste(input$category, "•", k, "-day MA"),
           x = NULL, y = y_lab) +
      theme_minimal(base_size = 12) +
      theme(legend.position = "right")
  })
}

shinyApp(ui, server)
```

Shiny app:

```{r}
# ============================================================================
# Municipal PV adoption in Germany — cumulative-registration map + daily chart
# ============================================================================
pacman::p_load(shiny, leaflet, sf, sfarrow, arrow, dplyr, ggplot2, here, tibble)

# ---- 0. data ---------------------------------------------------------------
base_dir <- here("8. shiny app", "8.1. shiny apps", "3")

# municipalities.parquet stores its geometry column as `geom`; federalState
# as `geometry`. Drop any empty geometries via name-agnostic base subsetting
# (st_make_valid can leave empties that leaflet::addPolygons cannot read).
muni_sf <- st_read_parquet(file.path(base_dir, "municipalities.parquet")) |>
  st_zm(drop = TRUE) |>
  st_transform(4326) |>
  st_make_valid() |>
  st_simplify(dTolerance = 0.0035, preserveTopology = TRUE) |>
  st_make_valid()
muni_sf <- muni_sf[!st_is_empty(muni_sf), ]

fs_sf <- st_read_parquet(file.path(base_dir, "federalState.parquet")) |>
  st_zm(drop = TRUE) |>
  st_transform(4326) |>
  st_make_valid()
fs_sf <- fs_sf[!st_is_empty(fs_sf), ]

mun_day <- read_parquet(file.path(base_dir, "mun_day.parquet")) |>
  mutate(date.registration.day = as.Date(date.registration.day))

# ---- 1. constants ----------------------------------------------------------
default_muni <- "08222000"                             # Mannheim

iran_milestones <- tibble::tribble(
  ~date,                 ~label,
  as.Date("2026-02-28"), "War begins / Hormuz closed",
  as.Date("2026-03-06"), "Ras Laffan gas hub shutdown",
  as.Date("2026-03-17"), "US naval campaign",
  as.Date("2026-03-21"), "US strikes Natanz",
  as.Date("2026-04-08"), "Ceasefire"
)

# colour scale: cumulative registrations per 100k over the whole sample.
# fixed bins so colours stay comparable when switching PV category.
cum_bins <- c(0, 50, 100, 150, 200, 300, 450, 700, Inf)

# ---- 2. UI -----------------------------------------------------------------
ui <- fluidPage(
  titlePanel("Iran war & municipal PV adoption — Germany"),
  fluidRow(
    column(4, radioButtons("category", "PV type:",
                           choices  = c("Balcony PVs", "Non-balcony PVs"),
                           selected = "Balcony PVs")),
    column(8, sliderInput("ma_window", "Moving-average window (days):",
                          min = 2, max = 30, value = 7, step = 1, width = "100%"))
  ),
  helpText(
    "Map colour = cumulative PV registrations per 100,000 inhabitants over the ",
    "whole sample (1 Jan – 7 May 2026). Grey = no data. ",
    "Click a municipality to load its daily series on the right."
  ),
  fluidRow(
    column(6, leafletOutput("map",   height = "640px")),
    column(6, plotOutput("chart",    height = "640px"))
  )
)

# ---- 3. server -------------------------------------------------------------
server <- function(input, output, session) {

  selected_muni <- reactiveVal(default_muni)

  # data for the chosen PV category
  cat_data <- reactive({
    mun_day |> filter(solar_category == input$category)
  })

  # cumulative registrations per 100k over the whole sample (per municipality)
  cum_tbl <- reactive({
    cat_data() |>
      group_by(MunicipalityCode) |>
      summarise(cum_per100k = sum(n) / first(population) * 1e5,
                .groups = "drop")
  })

  map_data <- reactive({
    muni_sf |> left_join(cum_tbl(), by = "MunicipalityCode")
  })

  # base map: tiles + federal-state borders (rendered once)
  output$map <- renderLeaflet({
    # default SVG renderer (not preferCanvas) — the canvas renderer fails to
    # clear hover labels on mouseout, leaving them stuck on the map.
    leaflet() |>
      addProviderTiles(providers$CartoDB.Positron) |>
      setView(lng = 10.45, lat = 51.16, zoom = 6) |>
      addMapPane("choropleth", zIndex = 410) |>
      addMapPane("borders",    zIndex = 420) |>
      addMapPane("highlight",  zIndex = 430) |>
      addPolygons(data = fs_sf, fill = FALSE, color = "#333333", weight = 1.6,
                  group = "states", options = pathOptions(pane = "borders"))
  })

  # choropleth: redrawn when the PV category changes
  observe({
    md  <- map_data()
    pal <- colorBin("YlGn", domain = md$cum_per100k, bins = cum_bins,
                    na.color = "#cfcfcf")

    leafletProxy("map") |>
      clearGroup("muni") |>
      clearControls() |>
      addPolygons(
        data        = md,
        group       = "muni",
        layerId     = ~MunicipalityCode,
        fillColor   = ~pal(cum_per100k),
        fillOpacity = 0.85,
        color       = "#ffffff",
        weight      = 0.25,
        options     = pathOptions(pane = "choropleth"),
        label       = ~ifelse(
          is.na(cum_per100k),
          paste0(Municipality, " — no data"),
          paste0(Municipality, ": ", round(cum_per100k), " per 100k")
        ),
        # bringToFront = FALSE: bringing a polygon to front on hover moves its
        # DOM node, so the browser drops the mouseout event and labels stick.
        highlightOptions = highlightOptions(weight = 1.5, color = "#000000",
                                            bringToFront = FALSE)
      ) |>
      addLegend(position = "bottomright", pal = pal, values = md$cum_per100k,
                title = "Cumulative reg.<br>per 100k inhab.",
                na.label = "no data")
  })

  # click a municipality -> update selection
  observeEvent(input$map_shape_click, {
    id <- input$map_shape_click$id
    if (!is.null(id)) selected_muni(id)
  })

  # highlight the selected municipality
  observe({
    sel <- muni_sf |> filter(MunicipalityCode == selected_muni())
    proxy <- leafletProxy("map") |> clearGroup("sel")
    if (nrow(sel) == 1) {
      proxy |> addPolygons(data = sel, group = "sel", fill = FALSE,
                           color = "#1f4e79", weight = 3,
                           options = pathOptions(pane = "highlight"))
    }
  })

  # per-municipality daily chart
  output$chart <- renderPlot({
    df <- cat_data() |>
      filter(MunicipalityCode == selected_muni()) |>
      arrange(date.registration.day)
    validate(need(nrow(df) > 0,
                  "No registration data for this municipality."))

    muni_name <- df$Municipality[1]
    has_pop   <- !any(is.na(df$population))
    df$y      <- if (has_pop) df$n_per100k else df$n

    k     <- input$ma_window
    df$ma <- as.numeric(stats::filter(df$y, rep(1 / k, k), sides = 1))

    y_lab <- if (has_pop) "Daily registrations per 100k inhab."
             else         "Daily registrations (raw count)"
    subt  <- paste0(input$category,
                    if (!has_pop)
                      "  •  no population data — showing raw counts" else "")

    y_top <- max(c(df$y, df$ma), na.rm = TRUE)
    ms    <- iran_milestones |>
      filter(date >= min(df$date.registration.day),
             date <= max(df$date.registration.day))

    ggplot(df, aes(x = date.registration.day)) +
      geom_col(aes(y = y), fill = "grey70", width = 1) +
      geom_vline(data = ms, aes(xintercept = date),
                 color = "black", linetype = "dashed", linewidth = 0.4) +
      geom_line(aes(y = ma), color = "#c0392b", linewidth = 1, na.rm = TRUE) +
      geom_text(data = ms, aes(x = date, y = y_top, label = label),
                angle = 90, hjust = 1, vjust = -0.3, size = 2.9,
                color = "grey20", check_overlap = TRUE) +
      scale_x_date(date_labels = "%b", date_breaks = "1 month") +
      scale_y_continuous(expand = expansion(mult = c(0, 0.20))) +
      labs(title    = paste0(muni_name, " — daily PV registrations"),
           subtitle = subt,
           x = NULL, y = y_lab,
           caption = paste0(k, "-day moving average (red). ",
                            "Black dashed lines = Iran-war milestones.")) +
      theme_minimal(base_size = 13)
  })
}

shinyApp(ui, server)
```