---
title: 'How to merge all files'
format: html
---



> Here I want to show how to merge all files once you download them from Market Master Register.



Libraries:

```{r}
pacman::p_load(xml2, dplyr, tidyr, purrr, readr)
```

Folder where you can find all files: *D:\1.b. ZEW\2. new\1. sequential.work\10. data for MastermarketRegister\May7_Gesamtdatenexport_20260507_26.1*

In the folder *1. data\1.1. all files* there are a few files named "EinheitenSolar_" + a number. I want to generate a unique dataset with all these files toegther.


Here you have the code to generate the big data:
```{r}


data_path <- "D:/1.b. ZEW/2. new/1. sequential.work/10. data for MastermarketRegister/May7_Gesamtdatenexport_20260507_26.1"

# List all files matching the pattern
files <- list.files(data_path, pattern = "EinheitenSolar", full.names = TRUE)
# files <- files[1:3]  # Uncomment to test with first 3 files only

# Initialize output
output <- NULL

# Loop through all files
for(i in 1:length(files)){

  # Define the XML file path
  xml_file <- files[i]

  # Read the XML file
  xml_data <- read_xml(xml_file)
  
  # Extract the root element (assuming all records are under a common parent node)
  root_node <- xml_find_all(xml_data, "/*/*")  # Adjust if needed based on XML structure
  
  # Convert XML to a generic data frame
  data_list <- map(root_node, ~ {
    # Extract all children nodes for the current record
    nodes <- xml_children(.x)
    # Create a named list with tag names as column names and their text values
    setNames(
      map_chr(nodes, xml_text),  # Extract text content for each node
      map_chr(nodes, xml_name)   # Extract the names of the nodes
    )
  })
  
  # Combine the list of records into a data frame
  data_frame <- bind_rows(data_list)
  
  # Append to output
  if(i == 1){
    output <- data_frame
  } else{
    output <- bind_rows(output, data_frame)
  }
  
  print(paste("Processed file", i, "of", length(files)))
}

# Display column names
print(names(output))


```
