---
title: 'Shiny app 1 — Balcony PVs across federal states'
format: html
---


> Map of balcony-PV adoption per federal state (population-adjusted) with a
> time slider, paired with a Germany-wide time series of balcony-PV
> registrations and vertical markers for the 2026 Iran-war milestones.

# 0. Libraries

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here,        # here()
  arrow,       # read_parquet()
  sfarrow,     # st_read_parquet()
  sf,          # st_transform(), st_as_sf()
  dplyr,       # rename, mutate, filter, left_join, group_by, summarise
  tibble,      # tribble()
  lubridate,   # floor_date()
  shiny,       # shinyApp(), fluidPage(), sliderInput()
  leaflet,     # leaflet(), addPolygons(), colorNumeric()
  ggplot2      # ggplot(), geom_line(), geom_vline()
)
```


# 1. Load and prepare the data

Per-frequency parquet files produced by
[4. datasets/4.2. federal state/2. report/3. dataset.qmd](../../../4.%20datasets/4.2.%20federal%20state/2.%20report/3.%20dataset.qmd).
Each file already contains `n`, `n_cumulative`, `n_per100k`,
`n_cumulative_per100k` (per 100 000 inhabitants, 2023 population). In this
chunk we also load the federal-state boundaries, define the Iran-war
milestones, and pre-compute the fixed colour-scale domain.

**Frequency choice.**

*Map (slider) — monthly.* The Iran-war milestones (28 Feb / 18 Mar / 8 Apr
2026) sit in three distinct months, so a monthly slider exposes the impact
one month at a time. Weekly would yield a very noisy choropleth (several
small states show 0–2 balcony registrations in any given week per 100 000
inhabitants), and quarterly collapses the entire war into one or two slider
positions.

*Time-series plot — weekly.* Aggregating to weekly is smooth enough to read
while still letting the three milestone dates fall into separate bars,
which would not be possible at monthly frequency.

```{r}
# --- federal-state PV data (weekly + monthly) -------------------------------
out_dir <- here("4. datasets", "4.2. federal state", "1. data", "3")

state_week  <- read_parquet(file.path(out_dir, "state_week.parquet"))
state_month <- read_parquet(file.path(out_dir, "state_month.parquet"))

state_week_b <- state_week |>
  filter(solar_category == "Balcony PVs") |>
  rename(date = date.registration.week)

state_month_b <- state_month |>
  filter(solar_category == "Balcony PVs") |>
  rename(date = date.registration.month)

# --- federal-state boundaries -----------------------------------------------
fs_map <- st_read_parquet(here("6. Maps", "6.1. Dataset", "federalState.parquet")) |>
  st_transform(4326)   # leaflet wants WGS84

# --- Iran-war milestones (vertical markers on the time-series plot) ---------
iran_milestones <- tibble::tribble(
  ~date,                  ~label,
  as.Date("2026-02-28"),  "War begins / Hormuz closed",
  as.Date("2026-03-18"),  "Major escalation",
  as.Date("2026-04-08"),  "Temporary 2-week ceasefire"
)

# --- fixed colour-scale domain so all slider positions are comparable -------
# one maximum per metric (per-period vs cumulative); the chosen one is used
# as the colour-scale domain so colours stay comparable across slider steps.
global_max_period <- max(state_month_b$n_per100k,            na.rm = TRUE)
global_max_cum    <- max(state_month_b$n_cumulative_per100k, na.rm = TRUE)
months_seq        <- sort(unique(state_month_b$date))
```


# 2. Shiny app

```{r}
ui <- fluidPage(
  titlePanel("Balcony-PV adoption across German federal states"),
  tags$p(
    style = "font-size: 13px; color: #555; margin-top: -8px;",
    tags$b("Note: "),
    "values shown on the map are balcony PV registrations ",
    tags$i("per 100 000 inhabitants"),
    " (2023 federal-state population). The colour scale is fixed across all ",
    "slider positions so periods are visually comparable."
  ),
  fluidRow(
    column(6,
      sliderInput("month_pick", "Month displayed on the map",
                  min       = min(months_seq),
                  max       = max(months_seq),
                  value     = max(months_seq),
                  timeFormat = "%b %Y",
                  step      = 31,
                  animate   = animationOptions(interval = 700),
                  width     = "100%"),
      radioButtons("map_mode", NULL,
                   choices  = c("Registrations in the selected month" = "period",
                                "Cumulative registrations up to the selected month" = "cum"),
                   selected = "period",
                   inline   = TRUE),
      leafletOutput("map", height = 560)
    ),
    column(6,
      tags$h4("Balcony PVs over time — one line per federal state (weekly)"),
      fluidRow(
        column(4,
          checkboxInput("show_lines",
                        "Show Iran-war milestone lines",
                        value = TRUE)
        ),
        column(8,
          sliderInput("ts_range", "Date range",
                      min   = min(state_week_b$date),
                      max   = max(state_week_b$date),
                      value = c(min(state_week_b$date),
                                max(state_week_b$date)),
                      timeFormat = "%b %Y",
                      width = "100%")
        )
      ),
      selectizeInput(
        "ts_states", "Federal states to display",
        choices  = sort(unique(state_week_b$FederalState.names)),
        selected = sort(unique(state_week_b$FederalState.names)),
        multiple = TRUE,
        width    = "100%",
        options  = list(
          placeholder      = "Type to search a federal state…",
          plugins          = list("remove_button"),
          closeAfterSelect = FALSE
        )
      ),
      plotOutput("ts", height = 560)
    )
  ),
  tags$div(
    style = "font-size: 12px; line-height: 1.4; margin-top: 10px;",
    tags$b("Iran-war 2026 milestones (vertical lines on the plot):"),
    tags$ul(
      style = "padding-left: 18px; margin-top: 4px;",
      tags$li(tags$b("28 Feb 2026"), " — War begins / Strait of Hormuz closed"),
      tags$li(tags$b("18 Mar 2026"), " — Major escalation"),
      tags$li(tags$b("08 Apr 2026"), " — Temporary 2-week ceasefire")
    )
  )
)

server <- function(input, output, session) {

  # snap the slider to the nearest available month in months_seq
  selected_month <- reactive({
    months_seq[ which.min(abs(months_seq - input$month_pick)) ]
  })

  # palette + scale specs that depend on the cumulative/per-period toggle.
  # Domain is fixed across all slider positions (per metric) so colours stay
  # comparable from one month to the next.
  map_spec <- reactive({
    if (input$map_mode == "cum") {
      list(
        col      = "n_cumulative_per100k",
        max      = global_max_cum,
        legend_t = "Cumulative balcony PVs<br/>per 100k pop.<br/>(up to month shown)",
        label_t  = "%.2f cumulative balcony PVs per 100k inhabitants"
      )
    } else {
      list(
        col      = "n_per100k",
        max      = global_max_period,
        legend_t = "Balcony PVs<br/>per 100k pop.<br/>(in month shown)",
        label_t  = "%.2f balcony PVs per 100k inhabitants (this month)"
      )
    }
  })

  pal <- reactive({
    colorNumeric("YlOrRd", domain = c(0, map_spec()$max), na.color = "#eeeeee")
  })

  # static base map (rendered once — tiles + bounds only)
  output$map <- renderLeaflet({
    leaflet(fs_map) |>
      addProviderTiles(providers$CartoDB.Positron) |>
      fitBounds(
        lng1 = as.numeric(st_bbox(fs_map)["xmin"]),
        lat1 = as.numeric(st_bbox(fs_map)["ymin"]),
        lng2 = as.numeric(st_bbox(fs_map)["xmax"]),
        lat2 = as.numeric(st_bbox(fs_map)["ymax"])
      )
  })

  # redraw the legend whenever the mode (= palette) changes
  observe({
    spec <- map_spec()
    leafletProxy("map") |>
      clearControls() |>
      addLegend(
        position = "bottomright",
        pal      = pal(),
        values   = c(0, spec$max),
        title    = spec$legend_t,
        opacity  = 0.9
      )
  })

  # update polygons whenever the slider or mode changes
  observe({
    m    <- selected_month()
    spec <- map_spec()

    slice_df <- state_month_b |>
      filter(date == m) |>
      select(FederalStateCode, all_of(spec$col))

    map_data <- fs_map |>
      left_join(slice_df, by = "FederalStateCode")

    values <- map_data[[spec$col]]

    labels <- sprintf(
      paste0("<b>%s</b><br/>", spec$label_t),
      map_data$FederalState,
      values
    ) |> lapply(htmltools::HTML)

    leafletProxy("map", data = map_data) |>
      clearShapes() |>
      addPolygons(
        fillColor   = pal()(values),
        weight      = 1,
        color       = "#444",
        fillOpacity = 0.85,
        label       = labels,
        highlightOptions = highlightOptions(
          weight = 2, color = "#000", fillOpacity = 0.95, bringToFront = TRUE
        )
      )
  })

  # weekly time series of balcony PVs per 100k — one line per federal state
  output$ts <- renderPlot({
    req(input$ts_states)

    plot_df <- state_week_b |>
      filter(FederalState.names %in% input$ts_states,
             date >= input$ts_range[1],
             date <= input$ts_range[2])

    p <- ggplot(plot_df,
                aes(date, n_per100k, colour = FederalState.names)) +
      geom_line(linewidth = 0.45, alpha = 0.85) +
      labs(x = NULL,
           y = "Balcony PVs per 100k inhabitants (weekly)",
           colour = "Federal state") +
      theme_minimal(base_size = 12) +
      theme(legend.position = "bottom") +
      guides(colour = guide_legend(nrow = 4, byrow = TRUE))

    if (isTRUE(input$show_lines)) {
      ml <- iran_milestones |>
        filter(date >= input$ts_range[1], date <= input$ts_range[2])
      if (nrow(ml) > 0) {
        p <- p +
          geom_vline(data = ml,
                     aes(xintercept = date),
                     colour = "black", linewidth = 0.5,
                     inherit.aes = FALSE) +
          geom_text(data = ml,
                    aes(x = date, y = Inf, label = label),
                    inherit.aes = FALSE,
                    hjust = 1.05, vjust = 1.2,
                    angle = 90, size = 3.2, fontface = "bold")
      }
    }

    p
  })
}

shinyApp(ui, server)
```
