---
title: 'Check the datasets'
format: html
---


> Here I want to check whether the datasets are correct or not.

**Claude chat**:
- CHECK (1)

Libraries:

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here,        # here()
  arrow,       # read_parquet()
  dplyr,       # glimpse(), filter(), arrange(), count()
  lubridate,   # year()
  ggplot2      # ggplot(), geom_col(), facet_wrap()
)
```

Datasets:
- *4. datasets\4.1. all germany\1. data\3*
- *4. datasets\4.2. federal state\1. data\3*


# 1. load the datasets
> Here I want to load the datasets.

Germany-level (one row per `solar_category` × date):

```{r}
germany_dir <- here("4. datasets", "4.1. all germany", "1. data", "3")

germany_day     <- read_parquet(file.path(germany_dir, "germany_day.parquet"))
germany_week    <- read_parquet(file.path(germany_dir, "germany_week.parquet"))
germany_month   <- read_parquet(file.path(germany_dir, "germany_month.parquet"))
germany_quarter <- read_parquet(file.path(germany_dir, "germany_quarter.parquet"))
germany_year    <- read_parquet(file.path(germany_dir, "germany_year.parquet"))
```

Federal-state level (one row per state × `solar_category` × date):

```{r}
state_dir <- here("4. datasets", "4.2. federal state", "1. data", "3")

state_day     <- read_parquet(file.path(state_dir, "state_day.parquet"))
state_week    <- read_parquet(file.path(state_dir, "state_week.parquet"))
state_month   <- read_parquet(file.path(state_dir, "state_month.parquet"))
state_quarter <- read_parquet(file.path(state_dir, "state_quarter.parquet"))
state_year    <- read_parquet(file.path(state_dir, "state_year.parquet"))
```

# 2. checks
> Here I want to check whether the datasest are correct or not.

## 2.1. Fast checks
> Here I want to perfomr basic checks.

Dimensions of all 10 files (expect more rows at finer frequency, and state files = 16 × Germany rows):

```{r}
sapply(
  list(
    germany_day     = germany_day,     germany_week  = germany_week,
    germany_month   = germany_month,   germany_quarter = germany_quarter,
    germany_year    = germany_year,
    state_day       = state_day,       state_week    = state_week,
    state_month     = state_month,     state_quarter = state_quarter,
    state_year      = state_year
  ),
  dim
)
```

Categories present (expect exactly `Balcony PVs` and `Non-balcony PVs`):

```{r}
unique(germany_day$solar_category)
unique(state_day$solar_category)
```

Federal-state coverage (expect 16 distinct names and 16 distinct NUTS codes):

```{r}
c(
  n_state_names = length(unique(state_day$FederalState.names)),
  n_state_codes = length(unique(state_day$FederalStateCode))
)
```

NA scan on numeric columns (expect all zeros):

```{r}
num_cols <- c("n", "n_cumulative", "n_per100k", "n_cumulative_per100k")

sapply(germany_day[, num_cols], function(x) sum(is.na(x)))
sapply(state_day[,   num_cols], function(x) sum(is.na(x)))
```

Glimpse the daily files (column names + types match expectations):

```{r}
glimpse(germany_day)
glimpse(state_day)
```

## 2.2. Plots
> Here I want to inspect the dataset with some plots.

I want to create a facet wrap plot for the federal states. For now, at yearly frequency.

```{r}
#| fig-width: 11
#| fig-height: 8

ggplot(state_year,
       aes(x = year(date.registration.year),
           y = n,
           colour = solar_category)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 1.2) +
  facet_wrap(~ FederalState.names, scales = "free_y", ncol = 4) +
  scale_colour_manual(values = c("Balcony PVs" = "#d95f02",
                                 "Non-balcony PVs" = "#1b9e77")) +
  labs(x = NULL,
       y = "Registrations per year",
       colour = NULL,
       title = "Yearly solar registrations by federal state",
       subtitle = "Free y-axis — shape of the series matters more than levels") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))
```

## 2.3. shiny app
> Here I want to make some checks with the shiny app.

Here I want to create a shiny app where you can select the following options:
- the frequency (from daily to yearly)
- option whether to dispaly the verticla lines for relevant dates of Iran war (somewhere place the dates)

The shiny app will show the facet wrap fo all federal states