---
title: 'Check the datasets'
format: html
---


> Here I want to check whether the datasets are correct or not.

**Claude chat**:
- CHECK (1)

Libraries:

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here,        # here()
  arrow,       # read_parquet()
  dplyr,       # glimpse(), filter(), arrange(), count()
  tibble,      # tribble()
  lubridate,   # year()
  ggplot2,     # ggplot(), geom_line(), facet_wrap()
  ggiraph,     # geom_line_interactive(), girafe()
  shiny        # shinyApp(), fluidPage(), sliderInput()
)
```

Datasets:
- *4. datasets\4.1. all germany\1. data\3*
- *4. datasets\4.2. federal state\1. data\3*


# 1. load the datasets
> Here I want to load the datasets.

Germany-level (one row per `solar_category` × date):

```{r}
germany_dir <- here("4. datasets", "4.1. all germany", "1. data", "3")

germany_day     <- read_parquet(file.path(germany_dir, "germany_day.parquet"))
germany_week    <- read_parquet(file.path(germany_dir, "germany_week.parquet"))
germany_month   <- read_parquet(file.path(germany_dir, "germany_month.parquet"))
germany_quarter <- read_parquet(file.path(germany_dir, "germany_quarter.parquet"))
germany_year    <- read_parquet(file.path(germany_dir, "germany_year.parquet"))
```

Federal-state level (one row per state × `solar_category` × date):

```{r}
state_dir <- here("4. datasets", "4.2. federal state", "1. data", "3")

state_day     <- read_parquet(file.path(state_dir, "state_day.parquet"))
state_week    <- read_parquet(file.path(state_dir, "state_week.parquet"))
state_month   <- read_parquet(file.path(state_dir, "state_month.parquet"))
state_quarter <- read_parquet(file.path(state_dir, "state_quarter.parquet"))
state_year    <- read_parquet(file.path(state_dir, "state_year.parquet"))
```

# 2. checks
> Here I want to check whether the datasest are correct or not.

## 2.1. Fast checks
> Here I want to perfomr basic checks.

Dimensions of all 10 files (expect more rows at finer frequency, and state files = 16 × Germany rows):

```{r}
sapply(
  list(
    germany_day     = germany_day,     germany_week  = germany_week,
    germany_month   = germany_month,   germany_quarter = germany_quarter,
    germany_year    = germany_year,
    state_day       = state_day,       state_week    = state_week,
    state_month     = state_month,     state_quarter = state_quarter,
    state_year      = state_year
  ),
  dim
)
```

Categories present (expect exactly `Balcony PVs` and `Non-balcony PVs`):

```{r}
unique(germany_day$solar_category)
unique(state_day$solar_category)
```

Federal-state coverage (expect 16 distinct names and 16 distinct NUTS codes):

```{r}
c(
  n_state_names = length(unique(state_day$FederalState.names)),
  n_state_codes = length(unique(state_day$FederalStateCode))
)
```

NA scan on numeric columns (expect all zeros):

```{r}
num_cols <- c("n", "n_cumulative", "n_per100k", "n_cumulative_per100k")

sapply(germany_day[, num_cols], function(x) sum(is.na(x)))
sapply(state_day[,   num_cols], function(x) sum(is.na(x)))
```

Glimpse the daily files (column names + types match expectations):

```{r}
glimpse(germany_day)
glimpse(state_day)
```

## 2.2. Plots
> Here I want to inspect the dataset with some plots.

I want to create a facet wrap plot for the federal states. For now, at yearly frequency.

```{r}
#| fig-width: 11
#| fig-height: 8

ggplot(state_year,
       aes(x = year(date.registration.year),
           y = n,
           colour = solar_category)) +
  geom_line(linewidth = 0.6) +
  geom_point(size = 1.2) +
  facet_wrap(~ FederalState.names, scales = "free_y", ncol = 4) +
  scale_colour_manual(values = c("Balcony PVs" = "#d95f02",
                                 "Non-balcony PVs" = "#1b9e77")) +
  labs(x = NULL,
       y = "Registrations per year",
       colour = NULL,
       title = "Yearly solar registrations by federal state",
       subtitle = "Free y-axis — shape of the series matters more than levels") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))
```

## 2.3. shiny app
> Here I want to make some checks with the shiny app.

Here I want to create a shiny app where you can select the following options:
- the frequency (from daily to yearly)
- option whether to dispaly the verticla lines for relevant dates of Iran war (somewhere place the dates)
- slider to decide the time range
- option to select whether to show balcony PVs, non-balcony PVs or both

Other things:
- always use the population adjusted series
- make the line interactive using ggiraph package
- for the daily frequency, show a MA of 7 days + daily series in the background in grey, less visibile

The shiny app will show the facet wrap fo all federal states.

Shiny app:

```{r}
# --- one row per frequency: which dataset to use + its date column ----------
freq_map <- list(
  Daily     = list(df = state_day,     date_col = "date.registration.day"),
  Weekly    = list(df = state_week,    date_col = "date.registration.week"),
  Monthly   = list(df = state_month,   date_col = "date.registration.month"),
  Quarterly = list(df = state_quarter, date_col = "date.registration.quarter"),
  Yearly    = list(df = state_year,    date_col = "date.registration.year")
)

iran_milestones <- tribble(
  ~date,                  ~label,
  as.Date("2026-02-28"),  "War begins / Hormuz closed",
  as.Date("2026-03-18"),  "Major escalation",
  as.Date("2026-04-08"),  "Temporary 2-week ceasefire"
)

# global date range (taken from the finest grain so the slider spans everything)
date_range_global <- range(state_day$date.registration.day, na.rm = TRUE)

ui <- fluidPage(
  titlePanel("Balcony / non-balcony PV registrations per 100k inhabitants — federal states"),
  sidebarLayout(
    sidebarPanel(
      width = 3,
      radioButtons("freq", "Frequency",
                   choices  = names(freq_map),
                   selected = "Monthly",
                   inline   = TRUE),
      sliderInput("dates", "Date range",
                  min        = date_range_global[1],
                  max        = date_range_global[2],
                  value      = date_range_global,
                  timeFormat = "%b %Y",
                  width      = "100%"),
      checkboxGroupInput("cats", "Solar category",
                         choices  = c("Balcony PVs", "Non-balcony PVs"),
                         selected = c("Balcony PVs", "Non-balcony PVs"),
                         inline   = TRUE),
      checkboxInput("show_lines",
                    "Show Iran-war milestone lines",
                    value = TRUE),
      tags$div(
        style = "font-size: 12px; line-height: 1.4; margin-top: 10px;",
        tags$b("Iran-war 2026 milestones:"),
        tags$ul(
          style = "padding-left: 18px; margin-top: 4px;",
          tags$li(tags$b("28 Feb 2026"), " — War begins / Hormuz closed"),
          tags$li(tags$b("18 Mar 2026"), " — Major escalation"),
          tags$li(tags$b("08 Apr 2026"), " — Temporary 2-week ceasefire")
        )
      )
    ),
    mainPanel(
      width = 9,
      girafeOutput("plot", height = "780px")
    )
  )
)

server <- function(input, output, session) {

  # pick the right per-frequency dataset and normalise its date column to `date`
  plot_df <- reactive({
    req(input$cats)
    spec <- freq_map[[input$freq]]
    df   <- spec$df
    df$date <- df[[spec$date_col]]
    df |>
      filter(solar_category %in% input$cats,
             date >= input$dates[1],
             date <= input$dates[2])
  })

  output$plot <- renderGirafe({
    df <- plot_df()
    req(nrow(df) > 0)

    is_daily <- input$freq == "Daily"

    # at daily frequency, compute a centred 7-day MA per (state × category)
    if (is_daily) {
      df <- df |>
        arrange(FederalState.names, solar_category, date) |>
        group_by(FederalState.names, solar_category) |>
        mutate(n_per100k_ma7 = as.numeric(
          stats::filter(n_per100k, rep(1 / 7, 7), sides = 2)
        )) |>
        ungroup()
    }

    p <- ggplot(df,
                aes(x = date,
                    colour = solar_category,
                    group  = solar_category)) +
      facet_wrap(~ FederalState.names, scales = "free_y", ncol = 4) +
      scale_colour_manual(values = c("Balcony PVs"     = "#d95f02",
                                     "Non-balcony PVs" = "#1b9e77")) +
      labs(x = NULL,
           y = "Registrations per 100k inhabitants",
           colour = NULL) +
      theme_minimal(base_size = 10) +
      theme(legend.position = "bottom",
            axis.text.x     = element_text(angle = 45, hjust = 1))

    if (is_daily) {
      # raw daily noise in the background (no tooltip) + 7-day MA on top
      p <- p +
        geom_line(aes(y = n_per100k),
                  colour     = "grey75",
                  linewidth  = 0.25,
                  alpha      = 0.7,
                  show.legend = FALSE) +
        geom_line_interactive(
          data = df |> filter(!is.na(n_per100k_ma7)),
          aes(y       = n_per100k_ma7,
              tooltip = sprintf("%s — %s\n%s: %.2f per 100k (7-day MA)",
                                FederalState.names,
                                format(date, "%Y-%m-%d"),
                                solar_category,
                                n_per100k_ma7),
              data_id = solar_category),
          linewidth = 0.6
        )
    } else {
      p <- p +
        geom_line_interactive(
          aes(y       = n_per100k,
              tooltip = sprintf("%s — %s\n%s: %.2f per 100k",
                                FederalState.names,
                                format(date, "%Y-%m-%d"),
                                solar_category,
                                n_per100k),
              data_id = solar_category),
          linewidth = 0.5
        ) +
        geom_point_interactive(
          aes(y       = n_per100k,
              tooltip = sprintf("%s — %s\n%s: %.2f per 100k",
                                FederalState.names,
                                format(date, "%Y-%m-%d"),
                                solar_category,
                                n_per100k),
              data_id = solar_category),
          size = 0.9
        )
    }

    if (isTRUE(input$show_lines)) {
      ml <- iran_milestones |>
        filter(date >= input$dates[1], date <= input$dates[2])
      if (nrow(ml) > 0) {
        p <- p +
          geom_vline(data = ml,
                     aes(xintercept = date),
                     colour    = "black",
                     linewidth = 0.4,
                     linetype  = "dashed",
                     inherit.aes = FALSE)
      }
    }

    girafe(
      ggobj      = p,
      width_svg  = 12,
      height_svg = 8,
      options    = list(
        opts_hover(css = "stroke-width:1.6;"),
        opts_tooltip(css = "background:#222;color:#fff;padding:4px 6px;border-radius:3px;font-size:11px;")
      )
    )
  })
}

shinyApp(ui, server)
```