---
title: 'Dataset'
format: html
---


> Here I want to show the dataset and how to create it.

# 1. dataset
> Dataset to use for the analysis.

Dataset:

```{r}

data <- open_dataset(here("01_data", "EinheitenSolar_all.parquet"))


# Step 1: rename German → English (explicit mapping, column order differs from old dataset)
data <- data |>
  rename(
    UnitMastrNumber                          = EinheitMastrNummer,
    DateLastUpdate                           = DatumLetzteAktualisierung,
    LocationMastrNumber                      = LokationMaStRNummer,
    GridOperatorVerificationStatus           = NetzbetreiberpruefungStatus,
    GridOperatorVerificationDate             = NetzbetreiberpruefungDatum,
    PlantOperatorMastrNumber                 = AnlagenbetreiberMastrNummer,
    Country                                  = Land,
    FederalState                             = Bundesland,
    District                                 = Landkreis,
    Municipality                             = Gemeinde,
    MunicipalityCode                         = Gemeindeschluessel,
    PostalCode                               = Postleitzahl,
    City                                     = Ort,
    RegistrationDate                         = Registrierungsdatum,
    CommissioningDate                        = Inbetriebnahmedatum,
    UnitSystemStatus                         = EinheitSystemstatus,
    UnitOperationalStatus                    = EinheitBetriebsstatus,
    NotPresentInMigratedUnits                = NichtVorhandenInMigriertenEinheiten,
    PowerGenerationUnitName                  = NameStromerzeugungseinheit,
    Weic_NotAvailable                        = Weic_nv,
    PowerPlantNumber_NotAvailable            = Kraftwerksnummer_nv,
    EnergySource                             = Energietraeger,
    GrossCapacity                            = Bruttoleistung,
    NetNominalCapacity                       = Nettonennleistung,
    RemoteControllabilityGridOperator        = FernsteuerbarkeitNb,
    FeedInType                               = Einspeisungsart,
    AssignedActivePowerInverter              = ZugeordneteWirkleistungWechselrichter,
    NumberOfModules                          = AnzahlModule,
    TypeOfSolarSystem                        = ArtDerSolaranlage,
    PowerLimitation                          = Leistungsbegrenzung,
    UniformOrientationAndTiltAngle           = EinheitlicheAusrichtungUndNeigungswinkel,
    MainOrientation                          = Hauptausrichtung,
    MainOrientationTiltAngle                 = HauptausrichtungNeigungswinkel,
    UsageArea                                = Nutzungsbereich,
    EegMastrNumber                           = EegMaStRNummer,
    StorageAtSameLocation                    = SpeicherAmGleichenOrt,
    SecondaryOrientation                     = Nebenausrichtung,
    SecondaryOrientationTiltAngle            = NebenausrichtungNeigungswinkel,
    Street                                   = Strasse,
    StreetNotFound                           = StrasseNichtGefunden,
    HouseNumber_NotAvailable                 = Hausnummer_nv,
    HouseNumberNotFound                      = HausnummerNichtGefunden,
    Longitude                                = Laengengrad,
    Latitude                                 = Breitengrad,
    DateStartTemporaryDecommissioning        = DatumBeginnVoruebergehendeStilllegung,
    HouseNumber                              = Hausnummer,
    AddressAddition                          = Adresszusatz,
    RemoteControllabilityDirectMarketer      = FernsteuerbarkeitDv,
    PreviousPlantOperatorMastrNumber         = AltAnlagenbetreiberMastrNummer,  # new variable
    DateOfOperatorChange                     = DatumDesBetreiberwechsels,
    DateRegistrationOfOperatorChange         = DatumRegistrierungDesBetreiberwechsels,
    CadastralDistrict                        = Gemarkung,
    FieldParcelNumbers                       = FlurFlurstuecknummern,
    ConnectionToExtraHighOrHighVoltage       = AnschlussAnHoechstOderHochSpannung,
    DatePermanentDecommissioning             = DatumEndgueltigeStilllegung,
    SizeOfUsedAreaInHectares                 = GroesseDerInAnspruchGenommenenFlaecheInHektar,
    PreviousUsageTypeOfArea                  = VorherigerNutzungsartenbereichDerFlaeche,
    PredominantUsageTypeBeforeSolarConstruction = UeberwiegendeNutzungsartDerFlaecheVorErrichtung,
    NameOfSolarPark                          = NameDesSolarparks,
    AdditionalFeaturesOfAreaAndCurrentUsage  = ZusaetzlicheMerkmaleDerFlaecheUndDerAktuellenFlaechennutzung,
    GenMastrNumber                           = GenMastrNummer,
    CommissioningDateAtCurrentLocation       = InbetriebnahmedatumAmAktuellenStandort,
    DispatchResponsible                      = Einsatzverantwortlicher,
    CitizenEnergy                            = Buergerenergie,
    DateResumptionOfOperation                = DatumWiederaufnahmeBetrieb,
    PlannedCommissioningDate                 = GeplantesInbetriebnahmedatum,
    ClearanceHeight                          = LichteHoehe,
    PowerPlantNumber                         = Kraftwerksnummer
  )

# Step 2: convert numeric columns (XML parsing stores everything as character)
data <- data |>
  mutate(
    TypeOfSolarSystem           = as.integer(TypeOfSolarSystem),
    Country                     = as.integer(Country),
    FederalState                = as.integer(FederalState),
    GrossCapacity               = as.numeric(GrossCapacity),
    NetNominalCapacity          = as.numeric(NetNominalCapacity),
    AssignedActivePowerInverter = as.numeric(AssignedActivePowerInverter),
    NumberOfModules             = as.integer(NumberOfModules)
  )

# Step 3: lookup tables
C1 <- data.frame(
  TypeOfSolarSystem.names = c(
    "Structural_facilities", "Other_Structural_facilities", "Open_space",
    "Bodies_water", "Large_parking_lot", "Plug_in_sola_system", "Onshore&Offshore_wind_power"
  ),
  TypeOfSolarSystem = as.integer(c(853, 2484, 852, 3002, 3058, 2961, NA))
)

C2 <- tibble(
  Country = as.integer(c(84, 231, 90, 66, 198, 206, 106)),
  Country.sn = c("Germany", "Switzerland", "Denmark", "Belgium", "Netherlands", "Austria", "France")
)

C3 <- data.frame(
  FederalState = as.integer(c(
    1409, 1402, 1400, 1403, 1412, 1408, 1410, 1407,
    1414, 1411, 1413, 1405, 1415, 1406, 1401, 1404
  )),
  FederalState.names = c(
    "North Rhine-Westphalia", "Baden-Württemberg", "Brandenburg", "Bavaria",
    "Saarland", "Lower Saxony", "Rhineland-Palatinate", "Mecklenburg-Western Pomerania",
    "Saxony-Anhalt", "Schleswig-Holstein", "Saxony", "Hesse",
    "Thuringia", "Hamburg", "Berlin", "Bremen"
  ),
  FederalStateCode = c(
    "DEA", "DE1", "DE4", "DE2",
    "DEC", "DE9", "DEB", "DEM",
    "DEE", "DEF", "DED", "DE7",
    "DEG", "DE6", "DE3", "DE5"
  )
)

# district code from map (no geometry needed)
districts_map <- st_read_parquet(here("1. ignore", "3. data", "2. maps", "districts.parquet")) |>
  st_drop_geometry() |>
  select(District, DistrictCode)

# Step 4: join lookups
data <- data |>
  left_join(C1, by = "TypeOfSolarSystem") |>
  relocate(TypeOfSolarSystem.names, .after = TypeOfSolarSystem) |>
  left_join(C2, by = "Country") |>
  relocate(Country.sn, .after = Country) |>
  left_join(C3, by = "FederalState") |>
  relocate(FederalState.names, .after = FederalState) |>
  relocate(FederalStateCode, .after = FederalState.names) |>
  left_join(districts_map, by = "District") |>
  relocate(DistrictCode, .after = District)

# Step 5: solar_category
reform_date <- as.Date("2024-05-16")

data <- data |>
  mutate(
    RegistrationDate  = as.Date(RegistrationDate),
    CommissioningDate = as.Date(CommissioningDate),
    solar_category = case_when(
      TypeOfSolarSystem == 2961 & RegistrationDate <  reform_date & AssignedActivePowerInverter <= 0.6 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 & RegistrationDate >= reform_date & AssignedActivePowerInverter <= 0.8 & GrossCapacity <= 2.0 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 ~ "Plug_in_sola_system - Non-Balcony",
      TRUE ~ TypeOfSolarSystem.names
    )
  )

# Step 6: filter Germany
data <- data |>
  filter(Country.sn == "Germany")

# Step 7: date variables (week, month, quarter, year)
data <- data |>
  mutate(
    date.registration.day     = RegistrationDate,
    date.registration.week    = floor_date(RegistrationDate,  "week"),
    date.registration.month   = floor_date(RegistrationDate,  "month"),
    date.registration.quarter = floor_date(RegistrationDate,  "quarter"),
    date.registration.year    = floor_date(RegistrationDate,  "year"),
    date.commission.day       = CommissioningDate,
    date.commission.week      = floor_date(CommissioningDate, "week"),
    date.commission.month     = floor_date(CommissioningDate, "month"),
    date.commission.quarter   = floor_date(CommissioningDate, "quarter"),
    date.commission.year      = floor_date(CommissioningDate, "year")
  )

# Step 8: fix MunicipalityCode
data <- data |>
  mutate(MunicipalityCode = str_pad(as.character(MunicipalityCode), 8, pad = "0"))

data
```


# 2. Dataset 2025-2026 (RegistrationDate)
> Here I want to show the code to create the final dataset (and using RegistrationDate to filter only for the years 2025 and 2026). Also, I want to add the day date variable (not only week, month, etc.). So the idea is to create a dataset where for eaach geographic level (Germany, Federal State, District, Municipality, Postal Code) and for each date variable (day, week, month, quarter, year) I have the number of PVs registered in each period. For example, if I select Municipality level and month date variable, I want to have a dataset with the number of PVs registered in each month for each municipality. I want to do this for all the combinations of geographic level and date variable.


```{r}
data <- open_dataset(here("01_data", "EinheitenSolar_all.parquet"))


# Step 1: rename German → English (explicit mapping, column order differs from old dataset)
data <- data |>
  rename(
    UnitMastrNumber                          = EinheitMastrNummer,
    DateLastUpdate                           = DatumLetzteAktualisierung,
    LocationMastrNumber                      = LokationMaStRNummer,
    GridOperatorVerificationStatus           = NetzbetreiberpruefungStatus,
    GridOperatorVerificationDate             = NetzbetreiberpruefungDatum,
    PlantOperatorMastrNumber                 = AnlagenbetreiberMastrNummer,
    Country                                  = Land,
    FederalState                             = Bundesland,
    District                                 = Landkreis,
    Municipality                             = Gemeinde,
    MunicipalityCode                         = Gemeindeschluessel,
    PostalCode                               = Postleitzahl,
    City                                     = Ort,
    RegistrationDate                         = Registrierungsdatum,
    CommissioningDate                        = Inbetriebnahmedatum,
    UnitSystemStatus                         = EinheitSystemstatus,
    UnitOperationalStatus                    = EinheitBetriebsstatus,
    NotPresentInMigratedUnits                = NichtVorhandenInMigriertenEinheiten,
    PowerGenerationUnitName                  = NameStromerzeugungseinheit,
    Weic_NotAvailable                        = Weic_nv,
    PowerPlantNumber_NotAvailable            = Kraftwerksnummer_nv,
    EnergySource                             = Energietraeger,
    GrossCapacity                            = Bruttoleistung,
    NetNominalCapacity                       = Nettonennleistung,
    RemoteControllabilityGridOperator        = FernsteuerbarkeitNb,
    FeedInType                               = Einspeisungsart,
    AssignedActivePowerInverter              = ZugeordneteWirkleistungWechselrichter,
    NumberOfModules                          = AnzahlModule,
    TypeOfSolarSystem                        = ArtDerSolaranlage,
    PowerLimitation                          = Leistungsbegrenzung,
    UniformOrientationAndTiltAngle           = EinheitlicheAusrichtungUndNeigungswinkel,
    MainOrientation                          = Hauptausrichtung,
    MainOrientationTiltAngle                 = HauptausrichtungNeigungswinkel,
    UsageArea                                = Nutzungsbereich,
    EegMastrNumber                           = EegMaStRNummer,
    StorageAtSameLocation                    = SpeicherAmGleichenOrt,
    SecondaryOrientation                     = Nebenausrichtung,
    SecondaryOrientationTiltAngle            = NebenausrichtungNeigungswinkel,
    Street                                   = Strasse,
    StreetNotFound                           = StrasseNichtGefunden,
    HouseNumber_NotAvailable                 = Hausnummer_nv,
    HouseNumberNotFound                      = HausnummerNichtGefunden,
    Longitude                                = Laengengrad,
    Latitude                                 = Breitengrad,
    DateStartTemporaryDecommissioning        = DatumBeginnVoruebergehendeStilllegung,
    HouseNumber                              = Hausnummer,
    AddressAddition                          = Adresszusatz,
    RemoteControllabilityDirectMarketer      = FernsteuerbarkeitDv,
    PreviousPlantOperatorMastrNumber         = AltAnlagenbetreiberMastrNummer,  # new variable
    DateOfOperatorChange                     = DatumDesBetreiberwechsels,
    DateRegistrationOfOperatorChange         = DatumRegistrierungDesBetreiberwechsels,
    CadastralDistrict                        = Gemarkung,
    FieldParcelNumbers                       = FlurFlurstuecknummern,
    ConnectionToExtraHighOrHighVoltage       = AnschlussAnHoechstOderHochSpannung,
    DatePermanentDecommissioning             = DatumEndgueltigeStilllegung,
    SizeOfUsedAreaInHectares                 = GroesseDerInAnspruchGenommenenFlaecheInHektar,
    PreviousUsageTypeOfArea                  = VorherigerNutzungsartenbereichDerFlaeche,
    PredominantUsageTypeBeforeSolarConstruction = UeberwiegendeNutzungsartDerFlaecheVorErrichtung,
    NameOfSolarPark                          = NameDesSolarparks,
    AdditionalFeaturesOfAreaAndCurrentUsage  = ZusaetzlicheMerkmaleDerFlaecheUndDerAktuellenFlaechennutzung,
    GenMastrNumber                           = GenMastrNummer,
    CommissioningDateAtCurrentLocation       = InbetriebnahmedatumAmAktuellenStandort,
    DispatchResponsible                      = Einsatzverantwortlicher,
    CitizenEnergy                            = Buergerenergie,
    DateResumptionOfOperation                = DatumWiederaufnahmeBetrieb,
    PlannedCommissioningDate                 = GeplantesInbetriebnahmedatum,
    ClearanceHeight                          = LichteHoehe,
    PowerPlantNumber                         = Kraftwerksnummer
  )

# Step 2: convert numeric columns (XML parsing stores everything as character)
data <- data |>
  mutate(
    TypeOfSolarSystem           = as.integer(TypeOfSolarSystem),
    Country                     = as.integer(Country),
    FederalState                = as.integer(FederalState),
    GrossCapacity               = as.numeric(GrossCapacity),
    NetNominalCapacity          = as.numeric(NetNominalCapacity),
    AssignedActivePowerInverter = as.numeric(AssignedActivePowerInverter),
    NumberOfModules             = as.integer(NumberOfModules)
  )

# Step 3: lookup tables
C1 <- data.frame(
  TypeOfSolarSystem.names = c(
    "Structural_facilities", "Other_Structural_facilities", "Open_space",
    "Bodies_water", "Large_parking_lot", "Plug_in_sola_system", "Onshore&Offshore_wind_power"
  ),
  TypeOfSolarSystem = as.integer(c(853, 2484, 852, 3002, 3058, 2961, NA))
)

C2 <- tibble(
  Country = as.integer(c(84, 231, 90, 66, 198, 206, 106)),
  Country.sn = c("Germany", "Switzerland", "Denmark", "Belgium", "Netherlands", "Austria", "France")
)

C3 <- data.frame(
  FederalState = as.integer(c(
    1409, 1402, 1400, 1403, 1412, 1408, 1410, 1407,
    1414, 1411, 1413, 1405, 1415, 1406, 1401, 1404
  )),
  FederalState.names = c(
    "North Rhine-Westphalia", "Baden-Württemberg", "Brandenburg", "Bavaria",
    "Saarland", "Lower Saxony", "Rhineland-Palatinate", "Mecklenburg-Western Pomerania",
    "Saxony-Anhalt", "Schleswig-Holstein", "Saxony", "Hesse",
    "Thuringia", "Hamburg", "Berlin", "Bremen"
  ),
  FederalStateCode = c(
    "DEA", "DE1", "DE4", "DE2",
    "DEC", "DE9", "DEB", "DEM",
    "DEE", "DEF", "DED", "DE7",
    "DEG", "DE6", "DE3", "DE5"
  )
)

# district code from map (no geometry needed)
districts_map <- st_read_parquet(here("1. ignore", "3. data", "2. maps", "districts.parquet")) |>
  st_drop_geometry() |>
  select(District, DistrictCode)

# Step 4: join lookups
data <- data |>
  left_join(C1, by = "TypeOfSolarSystem") |>
  relocate(TypeOfSolarSystem.names, .after = TypeOfSolarSystem) |>
  left_join(C2, by = "Country") |>
  relocate(Country.sn, .after = Country) |>
  left_join(C3, by = "FederalState") |>
  relocate(FederalState.names, .after = FederalState) |>
  relocate(FederalStateCode, .after = FederalState.names) |>
  left_join(districts_map, by = "District") |>
  relocate(DistrictCode, .after = District)

# Step 5: solar_category
reform_date <- as.Date("2024-05-16")

data <- data |>
  mutate(
    RegistrationDate  = as.Date(RegistrationDate),
    CommissioningDate = as.Date(CommissioningDate),
    solar_category = case_when(
      TypeOfSolarSystem == 2961 & RegistrationDate <  reform_date & AssignedActivePowerInverter <= 0.6 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 & RegistrationDate >= reform_date & AssignedActivePowerInverter <= 0.8 & GrossCapacity <= 2.0 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 ~ "Plug_in_sola_system - Non-Balcony",
      TRUE ~ TypeOfSolarSystem.names
    )
  )

# Step 6: filter Germany
data <- data |>
  filter(Country.sn == "Germany")

# Step 7: date variables (week, month, quarter, year)
data <- data |>
  mutate(
    date.registration.day     = RegistrationDate,
    date.registration.week    = floor_date(RegistrationDate,  "week"),
    date.registration.month   = floor_date(RegistrationDate,  "month"),
    date.registration.quarter = floor_date(RegistrationDate,  "quarter"),
    date.registration.year    = floor_date(RegistrationDate,  "year"),
    date.commission.day       = CommissioningDate,
    date.commission.week      = floor_date(CommissioningDate, "week"),
    date.commission.month     = floor_date(CommissioningDate, "month"),
    date.commission.quarter   = floor_date(CommissioningDate, "quarter"),
    date.commission.year      = floor_date(CommissioningDate, "year")
  )

# Step 8: fix MunicipalityCode
data <- data |>
  mutate(MunicipalityCode = str_pad(as.character(MunicipalityCode), 8, pad = "0"))


# Add filter for 2025 and 2026
data <- data |>
  filter(RegistrationDate >= as.Date("2025-01-01"),
         RegistrationDate <= as.Date("2026-12-31"),
         is.na(CommissioningDate) | (CommissioningDate >= as.Date("2025-01-01") & CommissioningDate <= as.Date("2026-12-31")))

```


check:

```{r}
# districts in data with no match in map (DistrictCode is NA)
data |>
  filter(is.na(DistrictCode)) |>
  distinct(District) |>
  collect()

# names in data but not in map
data_districts <- data |> distinct(District) |> collect() |> arrange(District)
map_districts  <- districts_map |> distinct(District) |> arrange(District)

setdiff(data_districts$District, map_districts$District)  # in data, not in map
setdiff(map_districts$District,  data_districts$District) # in map, not in data
```

## 2.1 All Germany
Here I want to create a dataset for the whole Germany, at all frequencies (day, week, month and quarter), using the Registration date variable (not commission date):

```{r}
germany_day <- data |>
  group_by(solar_category, date.registration.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.day = seq(min(date.registration.day), max(date.registration.day), by = "day"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.day) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_week <- data |>
  group_by(solar_category, date.registration.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.week = seq(min(date.registration.week), max(date.registration.week), by = "week"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.week) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_month <- data |>
  group_by(solar_category, date.registration.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.month = seq(min(date.registration.month), max(date.registration.month), by = "month"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.month) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_quarter <- data |>
  group_by(solar_category, date.registration.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.quarter = seq(min(date.registration.quarter), max(date.registration.quarter), by = "quarter"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.quarter) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

## 2.2 Federal States

```{r}
state_day <- data |>
  group_by(FederalState.names, solar_category, date.registration.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(FederalState.names, solar_category, date.registration.day = seq(min(date.registration.day), max(date.registration.day), by = "day"), fill = list(n = 0)) |>
  arrange(FederalState.names, solar_category, date.registration.day) |>
  group_by(FederalState.names, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

state_week <- data |>
  group_by(FederalState.names, solar_category, date.registration.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(FederalState.names, solar_category, date.registration.week = seq(min(date.registration.week), max(date.registration.week), by = "week"), fill = list(n = 0)) |>
  arrange(FederalState.names, solar_category, date.registration.week) |>
  group_by(FederalState.names, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

state_month <- data |>
  group_by(FederalState.names, solar_category, date.registration.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(FederalState.names, solar_category, date.registration.month = seq(min(date.registration.month), max(date.registration.month), by = "month"), fill = list(n = 0)) |>
  arrange(FederalState.names, solar_category, date.registration.month) |>
  group_by(FederalState.names, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

state_quarter <- data |>
  group_by(FederalState.names, solar_category, date.registration.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(FederalState.names, solar_category, date.registration.quarter = seq(min(date.registration.quarter), max(date.registration.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(FederalState.names, solar_category, date.registration.quarter) |>
  group_by(FederalState.names, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

## 2.3 Districts

```{r}
district_day <- data |>
  group_by(District, solar_category, date.registration.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(District, solar_category, date.registration.day = seq(min(date.registration.day), max(date.registration.day), by = "day"), fill = list(n = 0)) |>
  arrange(District, solar_category, date.registration.day) |>
  group_by(District, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

district_week <- data |>
  group_by(District, solar_category, date.registration.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(District, solar_category, date.registration.week = seq(min(date.registration.week), max(date.registration.week), by = "week"), fill = list(n = 0)) |>
  arrange(District, solar_category, date.registration.week) |>
  group_by(District, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

district_month <- data |>
  group_by(District, solar_category, date.registration.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(District, solar_category, date.registration.month = seq(min(date.registration.month), max(date.registration.month), by = "month"), fill = list(n = 0)) |>
  arrange(District, solar_category, date.registration.month) |>
  group_by(District, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

district_quarter <- data |>
  group_by(District, solar_category, date.registration.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(District, solar_category, date.registration.quarter = seq(min(date.registration.quarter), max(date.registration.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(District, solar_category, date.registration.quarter) |>
  group_by(District, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

## 2.4 Municipalities

```{r}
mun_day <- data |>
  group_by(Municipality, MunicipalityCode, solar_category, date.registration.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(nesting(Municipality, MunicipalityCode), solar_category, date.registration.day = seq(min(date.registration.day), max(date.registration.day), by = "day"), fill = list(n = 0)) |>
  arrange(Municipality, solar_category, date.registration.day) |>
  group_by(Municipality, MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_week <- data |>
  group_by(Municipality, MunicipalityCode, solar_category, date.registration.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(nesting(Municipality, MunicipalityCode), solar_category, date.registration.week = seq(min(date.registration.week), max(date.registration.week), by = "week"), fill = list(n = 0)) |>
  arrange(Municipality, solar_category, date.registration.week) |>
  group_by(Municipality, MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_month <- data |>
  group_by(Municipality, MunicipalityCode, solar_category, date.registration.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(nesting(Municipality, MunicipalityCode), solar_category, date.registration.month = seq(min(date.registration.month), max(date.registration.month), by = "month"), fill = list(n = 0)) |>
  arrange(Municipality, solar_category, date.registration.month) |>
  group_by(Municipality, MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_quarter <- data |>
  group_by(Municipality, MunicipalityCode, solar_category, date.registration.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(nesting(Municipality, MunicipalityCode), solar_category, date.registration.quarter = seq(min(date.registration.quarter), max(date.registration.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(Municipality, solar_category, date.registration.quarter) |>
  group_by(Municipality, MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

## 2.5 Postal Codes

```{r}
plz_day <- data |>
  group_by(PostalCode, solar_category, date.registration.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(PostalCode, solar_category, date.registration.day = seq(min(date.registration.day), max(date.registration.day), by = "day"), fill = list(n = 0)) |>
  arrange(PostalCode, solar_category, date.registration.day) |>
  group_by(PostalCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

plz_week <- data |>
  group_by(PostalCode, solar_category, date.registration.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(PostalCode, solar_category, date.registration.week = seq(min(date.registration.week), max(date.registration.week), by = "week"), fill = list(n = 0)) |>
  arrange(PostalCode, solar_category, date.registration.week) |>
  group_by(PostalCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

plz_month <- data |>
  group_by(PostalCode, solar_category, date.registration.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(PostalCode, solar_category, date.registration.month = seq(min(date.registration.month), max(date.registration.month), by = "month"), fill = list(n = 0)) |>
  arrange(PostalCode, solar_category, date.registration.month) |>
  group_by(PostalCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

plz_quarter <- data |>
  group_by(PostalCode, solar_category, date.registration.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(PostalCode, solar_category, date.registration.quarter = seq(min(date.registration.quarter), max(date.registration.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(PostalCode, solar_category, date.registration.quarter) |>
  group_by(PostalCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```


# 3. Dataset 2025-2026 (CommissionDate)

## 3.1 All Germany

```{r}
germany_com_day <- data |>
  filter(!is.na(date.commission.day)) |>
  group_by(solar_category, date.commission.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.commission.day = seq(min(date.commission.day), max(date.commission.day), by = "day"), fill = list(n = 0)) |>
  arrange(solar_category, date.commission.day) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_com_week <- data |>
  filter(!is.na(date.commission.week)) |>
  group_by(solar_category, date.commission.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.commission.week = seq(min(date.commission.week), max(date.commission.week), by = "week"), fill = list(n = 0)) |>
  arrange(solar_category, date.commission.week) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_com_month <- data |>
  filter(!is.na(date.commission.month)) |>
  group_by(solar_category, date.commission.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.commission.month = seq(min(date.commission.month), max(date.commission.month), by = "month"), fill = list(n = 0)) |>
  arrange(solar_category, date.commission.month) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_com_quarter <- data |>
  filter(!is.na(date.commission.quarter)) |>
  group_by(solar_category, date.commission.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.commission.quarter = seq(min(date.commission.quarter), max(date.commission.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(solar_category, date.commission.quarter) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

## 3.2 Federal States

```{r}
state_com_day <- data |>
  filter(!is.na(date.commission.day)) |>
  group_by(FederalState.names, solar_category, date.commission.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(FederalState.names, solar_category, date.commission.day = seq(min(date.commission.day), max(date.commission.day), by = "day"), fill = list(n = 0)) |>
  arrange(FederalState.names, solar_category, date.commission.day) |>
  group_by(FederalState.names, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

state_com_week <- data |>
  filter(!is.na(date.commission.week)) |>
  group_by(FederalState.names, solar_category, date.commission.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(FederalState.names, solar_category, date.commission.week = seq(min(date.commission.week), max(date.commission.week), by = "week"), fill = list(n = 0)) |>
  arrange(FederalState.names, solar_category, date.commission.week) |>
  group_by(FederalState.names, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

state_com_month <- data |>
  filter(!is.na(date.commission.month)) |>
  group_by(FederalState.names, solar_category, date.commission.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(FederalState.names, solar_category, date.commission.month = seq(min(date.commission.month), max(date.commission.month), by = "month"), fill = list(n = 0)) |>
  arrange(FederalState.names, solar_category, date.commission.month) |>
  group_by(FederalState.names, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

state_com_quarter <- data |>
  filter(!is.na(date.commission.quarter)) |>
  group_by(FederalState.names, solar_category, date.commission.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(FederalState.names, solar_category, date.commission.quarter = seq(min(date.commission.quarter), max(date.commission.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(FederalState.names, solar_category, date.commission.quarter) |>
  group_by(FederalState.names, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

## 3.3 Districts

```{r}
district_com_day <- data |>
  filter(!is.na(date.commission.day)) |>
  group_by(District, solar_category, date.commission.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(District, solar_category, date.commission.day = seq(min(date.commission.day), max(date.commission.day), by = "day"), fill = list(n = 0)) |>
  arrange(District, solar_category, date.commission.day) |>
  group_by(District, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

district_com_week <- data |>
  filter(!is.na(date.commission.week)) |>
  group_by(District, solar_category, date.commission.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(District, solar_category, date.commission.week = seq(min(date.commission.week), max(date.commission.week), by = "week"), fill = list(n = 0)) |>
  arrange(District, solar_category, date.commission.week) |>
  group_by(District, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

district_com_month <- data |>
  filter(!is.na(date.commission.month)) |>
  group_by(District, solar_category, date.commission.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(District, solar_category, date.commission.month = seq(min(date.commission.month), max(date.commission.month), by = "month"), fill = list(n = 0)) |>
  arrange(District, solar_category, date.commission.month) |>
  group_by(District, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

district_com_quarter <- data |>
  filter(!is.na(date.commission.quarter)) |>
  group_by(District, solar_category, date.commission.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(District, solar_category, date.commission.quarter = seq(min(date.commission.quarter), max(date.commission.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(District, solar_category, date.commission.quarter) |>
  group_by(District, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

## 3.4 Municipalities

```{r}
mun_com_day <- data |>
  filter(!is.na(date.commission.day)) |>
  group_by(Municipality, MunicipalityCode, solar_category, date.commission.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(nesting(Municipality, MunicipalityCode), solar_category, date.commission.day = seq(min(date.commission.day), max(date.commission.day), by = "day"), fill = list(n = 0)) |>
  arrange(Municipality, solar_category, date.commission.day) |>
  group_by(Municipality, MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_com_week <- data |>
  filter(!is.na(date.commission.week)) |>
  group_by(Municipality, MunicipalityCode, solar_category, date.commission.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(nesting(Municipality, MunicipalityCode), solar_category, date.commission.week = seq(min(date.commission.week), max(date.commission.week), by = "week"), fill = list(n = 0)) |>
  arrange(Municipality, solar_category, date.commission.week) |>
  group_by(Municipality, MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_com_month <- data |>
  filter(!is.na(date.commission.month)) |>
  group_by(Municipality, MunicipalityCode, solar_category, date.commission.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(nesting(Municipality, MunicipalityCode), solar_category, date.commission.month = seq(min(date.commission.month), max(date.commission.month), by = "month"), fill = list(n = 0)) |>
  arrange(Municipality, solar_category, date.commission.month) |>
  group_by(Municipality, MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

mun_com_quarter <- data |>
  filter(!is.na(date.commission.quarter)) |>
  group_by(Municipality, MunicipalityCode, solar_category, date.commission.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(nesting(Municipality, MunicipalityCode), solar_category, date.commission.quarter = seq(min(date.commission.quarter), max(date.commission.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(Municipality, solar_category, date.commission.quarter) |>
  group_by(Municipality, MunicipalityCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

## 3.5 Postal Codes

```{r}
plz_com_day <- data |>
  filter(!is.na(date.commission.day)) |>
  group_by(PostalCode, solar_category, date.commission.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(PostalCode, solar_category, date.commission.day = seq(min(date.commission.day), max(date.commission.day), by = "day"), fill = list(n = 0)) |>
  arrange(PostalCode, solar_category, date.commission.day) |>
  group_by(PostalCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

plz_com_week <- data |>
  filter(!is.na(date.commission.week)) |>
  group_by(PostalCode, solar_category, date.commission.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(PostalCode, solar_category, date.commission.week = seq(min(date.commission.week), max(date.commission.week), by = "week"), fill = list(n = 0)) |>
  arrange(PostalCode, solar_category, date.commission.week) |>
  group_by(PostalCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

plz_com_month <- data |>
  filter(!is.na(date.commission.month)) |>
  group_by(PostalCode, solar_category, date.commission.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(PostalCode, solar_category, date.commission.month = seq(min(date.commission.month), max(date.commission.month), by = "month"), fill = list(n = 0)) |>
  arrange(PostalCode, solar_category, date.commission.month) |>
  group_by(PostalCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

plz_com_quarter <- data |>
  filter(!is.na(date.commission.quarter)) |>
  group_by(PostalCode, solar_category, date.commission.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(PostalCode, solar_category, date.commission.quarter = seq(min(date.commission.quarter), max(date.commission.quarter), by = "quarter"), fill = list(n = 0)) |>
  arrange(PostalCode, solar_category, date.commission.quarter) |>
  group_by(PostalCode, solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```

# 4. General function
> Here I want to create a function where you can select sone inputs and you get a dataset. Inputs:
- date variable (RegistrationDate or CommissioningDate)
- date frequency (day, week, month, quarter, year)
- geographic level (Germany, Federal State, District, Municipality, Postal Code)

I want to dispaly the dataset only for 2 years of data, i.e., 2025 and 2026. So, for example, if I select RegistrationDate, month frequency and Municipality level, I want to get a dataset with the number of PVs registered in each month for each municipality, but only for the years 2025 and 2026.

## 4.1 Function (with all code)
> Here I want to show the whole code to understand the structure.

```{r}
pv_dataset <- function(date_var  = "RegistrationDate",   # "RegistrationDate" or "CommissioningDate"
                       frequency = "month",              # "day", "week", "month", "quarter", "year"
                       geo_level = "Germany") {          # "Germany", "FederalState", "District", "Municipality", "PostalCode"

  # --- prepare data ---
  data <- open_dataset(here("01_data", "EinheitenSolar_all.parquet")) |>
    rename(
      RegistrationDate            = Registrierungsdatum,
      CommissioningDate           = Inbetriebnahmedatum,
      FederalState.names          = Bundesland,
      District                    = Landkreis,
      Municipality                = Gemeinde,
      MunicipalityCode            = Gemeindeschluessel,
      PostalCode                  = Postleitzahl,
      TypeOfSolarSystem           = ArtDerSolaranlage,
      GrossCapacity               = Bruttoleistung,
      AssignedActivePowerInverter = ZugeordneteWirkleistungWechselrichter,
      Country                     = Land
    ) |>
    mutate(
      RegistrationDate            = as.Date(RegistrationDate),
      CommissioningDate           = as.Date(CommissioningDate),
      TypeOfSolarSystem           = as.integer(TypeOfSolarSystem),
      Country                     = as.integer(Country),
      GrossCapacity               = as.numeric(GrossCapacity),
      AssignedActivePowerInverter = as.numeric(AssignedActivePowerInverter)
    ) |>
    filter(Country == 84L) |>
    mutate(
      solar_category = case_when(
        TypeOfSolarSystem == 2961L & RegistrationDate <  as.Date("2024-05-16") & AssignedActivePowerInverter <= 0.6                      ~ "Plug_in_sola_system - Balcony",
        TypeOfSolarSystem == 2961L & RegistrationDate >= as.Date("2024-05-16") & AssignedActivePowerInverter <= 0.8 & GrossCapacity <= 2.0 ~ "Plug_in_sola_system - Balcony",
        TypeOfSolarSystem == 2961L  ~ "Plug_in_sola_system - Non-Balcony",
        TypeOfSolarSystem == 853L   ~ "Structural_facilities",
        TypeOfSolarSystem == 2484L  ~ "Other_Structural_facilities",
        TypeOfSolarSystem == 852L   ~ "Open_space",
        TypeOfSolarSystem == 3002L  ~ "Bodies_water",
        TypeOfSolarSystem == 3058L  ~ "Large_parking_lot",
        TRUE                        ~ "Other"
      ),
      date.registration.day     = RegistrationDate,
      date.registration.week    = floor_date(RegistrationDate,  "week"),
      date.registration.month   = floor_date(RegistrationDate,  "month"),
      date.registration.quarter = floor_date(RegistrationDate,  "quarter"),
      date.registration.year    = floor_date(RegistrationDate,  "year"),
      date.commission.day       = CommissioningDate,
      date.commission.week      = floor_date(CommissioningDate, "week"),
      date.commission.month     = floor_date(CommissioningDate, "month"),
      date.commission.quarter   = floor_date(CommissioningDate, "quarter"),
      date.commission.year      = floor_date(CommissioningDate, "year")
    ) |>
    filter(RegistrationDate >= as.Date("2025-01-01"),
           RegistrationDate <= as.Date("2026-12-31"))

  # --- date column and geo columns ---
  date_prefix <- if (date_var == "RegistrationDate") "date.registration" else "date.commission"
  date_col    <- paste0(date_prefix, ".", frequency)

  geo_cols <- switch(geo_level,
    "Germany"      = character(0),
    "FederalState" = "FederalState.names",
    "District"     = "District",
    "Municipality" = c("Municipality", "MunicipalityCode"),
    "PostalCode"   = "PostalCode"
  )

  group_cols <- c(geo_cols, "solar_category", date_col)

  # --- aggregate ---
  result <- data |>
    group_by(across(all_of(group_cols))) |>
    summarise(n = n(), .groups = "drop") |>
    collect()

  # fill zeros
  date_seq <- seq(min(result[[date_col]]), max(result[[date_col]]), by = frequency)

  if (length(geo_cols) == 0) {
    result <- result |>
      complete(solar_category, !!sym(date_col) := date_seq, fill = list(n = 0))
  } else if (length(geo_cols) == 1) {
    result <- result |>
      complete(.data[[geo_cols]], solar_category, !!sym(date_col) := date_seq, fill = list(n = 0))
  } else {
    result <- result |>
      complete(nesting(!!!syms(geo_cols)), solar_category, !!sym(date_col) := date_seq, fill = list(n = 0))
  }

  # cumulative sum
  result <- result |>
    arrange(across(all_of(c(geo_cols, "solar_category", date_col)))) |>
    group_by(across(all_of(c(geo_cols, "solar_category")))) |>
    mutate(n_cumulative = cumsum(n)) |>
    ungroup()

  return(result)
}

# examples
pv_dataset(date_var = "RegistrationDate",  frequency = "month",   geo_level = "Germany")
pv_dataset(date_var = "RegistrationDate",  frequency = "month",   geo_level = "Municipality")
pv_dataset(date_var = "CommissioningDate", frequency = "quarter", geo_level = "FederalState")
```

## 4.2 Function (load saved datasets)
> Here I want to create the function that it gives you the dataset but instead of creating the dataset, it just load the files in the folder (*19. OTHER TASKS\1. Iran\3. data\1. dataset*). It has the same inputs, but instead of running the code, it just load the corresponding file. For example, if I select RegistrationDate, month frequency and Municipality level, it will load the file *mun_month.parquet* from the folder.


```{r}
pv_load <- function(date_var  = "RegistrationDate",   # "RegistrationDate" or "CommissioningDate"
                    frequency = "month",              # "day", "week", "month", "quarter"
                    geo_level = "Germany") {          # "Germany", "FederalState", "District", "Municipality", "PostalCode"

  # date folder
  date_folder <- switch(date_var,
    "RegistrationDate"  = "1. RegsitrationDate",
    "CommissioningDate" = "2. CommissionDate"
  )

  # geo subfolder
  geo_folder <- switch(geo_level,
    "Germany"      = "5. Germany",
    "FederalState" = "1. Federal State",
    "District"     = "2. District",
    "Municipality" = "3. Municipality",
    "PostalCode"   = "4. Postal Code"
  )

  # file prefix
  com_suffix <- if (date_var == "CommissioningDate") "_com" else ""

  prefix <- switch(geo_level,
    "Germany"      = paste0("germany",  com_suffix),
    "FederalState" = paste0("state",    com_suffix),
    "District"     = paste0("district", com_suffix),
    "Municipality" = paste0("mun",      com_suffix),
    "PostalCode"   = paste0("plz",      com_suffix)
  )

  file_name <- paste0(prefix, "_", frequency, ".parquet")

  read_parquet(here("1. ignore", "3. data", "1. dataset",
                    date_folder, geo_folder, file_name))
}

# examples
pv_load(date_var = "RegistrationDate",  frequency = "month",   geo_level = "Municipality")
pv_load(date_var = "CommissioningDate", frequency = "quarter", geo_level = "FederalState")
```

Load datasets separately:

```{r}
# via function
df_fn_1 <- pv_load(date_var = "RegistrationDate",  frequency = "month",   geo_level = "Municipality")
df_fn_2 <- pv_load(date_var = "CommissioningDate", frequency = "quarter", geo_level = "FederalState")

# manually
df_man_1 <- read_parquet(here("1. ignore", "3. data", "1. dataset",
                               "1. RegsitrationDate", "3. Municipality", "mun_month.parquet"))
df_man_2 <- read_parquet(here("1. ignore", "3. data", "1. dataset",
                               "2. CommissionDate",   "1. Federal State", "state_com_quarter.parquet"))
```

Check if they are identical:

```{r}
identical(df_fn_1, df_man_1)
identical(df_fn_2, df_man_2)
```