---
title: 'Heidelberg data: analysis'
format: html
---



# 1. Load the dataset
> Here I load the datasets and I filter for Heidelberg municipality (MunicipalityCode = 08221000)

Load the dataset using the function that admits the following inputs:

- *date_var*: `"RegistrationDate"` or `"CommissioningDate"`
- *frequency*: `"day"`, `"week"`, `"month"`, or `"quarter"`
- *geo_level*: `"Germany"`, `"FederalState"`, `"District"`, `"Municipality"`, or `"PostalCode"`


```{r}
pv_load <- function(date_var  = "RegistrationDate",   # "RegistrationDate" or "CommissioningDate"
                    frequency = "month",              # "day", "week", "month", "quarter"
                    geo_level = "Germany") {          # "Germany", "FederalState", "District", "Municipality", "PostalCode"

  # date folder
  date_folder <- switch(date_var,
    "RegistrationDate"  = "1. RegsitrationDate",
    "CommissioningDate" = "2. CommissionDate"
  )

  # geo subfolder
  geo_folder <- switch(geo_level,
    "Germany"      = "5. Germany",
    "FederalState" = "1. Federal State",
    "District"     = "2. District",
    "Municipality" = "3. Municipality",
    "PostalCode"   = "4. Postal Code"
  )

  # file prefix
  com_suffix <- if (date_var == "CommissioningDate") "_com" else ""

  prefix <- switch(geo_level,
    "Germany"      = paste0("germany",  com_suffix),
    "FederalState" = paste0("state",    com_suffix),
    "District"     = paste0("district", com_suffix),
    "Municipality" = paste0("mun",      com_suffix),
    "PostalCode"   = paste0("plz",      com_suffix)
  )

  file_name <- paste0(prefix, "_", frequency, ".parquet")

  read_parquet(here("01_data", "1. dataset",
                    date_folder, geo_folder, file_name))
}

# examples
pv_load(date_var = "RegistrationDate",  frequency = "month",   geo_level = "Municipality")
pv_load(date_var = "CommissioningDate", frequency = "quarter", geo_level = "FederalState")

```

Now, I want to filter for Heidelberg municipality:

```{r}
h1.day <- pv_load(date_var = "RegistrationDate",
            frequency = "day",
            geo_level = "Municipality") |>
  filter(MunicipalityCode == "08221000")

h1.week <- pv_load(date_var = "RegistrationDate",
            frequency = "week",
            geo_level = "Municipality") |>
  filter(MunicipalityCode == "08221000")

h1.month <- pv_load(date_var = "RegistrationDate",
            frequency = "month",
            geo_level = "Municipality") |>
  filter(MunicipalityCode == "08221000")
```

# 2. Iran war: some facts
> Here I want to show the facts relevant for the Iran War.

Table with most important dates:

```{r}
hormuz_dates <- data.frame(
  Date = c(
    "28 Feb – 1 Mar 2026",
    "2 Mar 2026",
    "6 Mar 2026",
    "~10 Mar 2026",
    "~17 Mar 2026",
    "21 Mar 2026",
    "8 Apr 2026"
  ),
  Event = c(
    "Iran closes the Strait of Hormuz; shipping grinds to standstill — 150+ oil tankers and freight ships stalled",
    "Iran formally declares no oil will leave the area; Iran seizes two oil tankers passing through the strait",
    "Satellite imagery confirms Ras Laffan (Qatar's main gas export hub) shutdown — fuel prices spike",
    "GPS jamming near the strait disrupts tanker navigation, raising accident risk",
    "US launches naval campaign to reopen the strait; Iran warns Bahrain and UAE they 'will be the first to lose' if they join",
    "US strikes Natanz nuclear facility with bunker busters; Iran vows to keep strait closed",
    "Ceasefire agreed — strait partially reopens; Iran retains toll collection in Chinese yuan"
  )
)

knitr::kable(hormuz_dates, col.names = c("Date", "Event"))
```

# 3. Solar Categories over time
> Here for each frequency I want to see the general evolution of the solar units.

## 3.1 Evolution Solar Categories: daily frequency

Dataset:

```{r}
h1.day <- pv_load(date_var = "RegistrationDate",
            frequency = "day",
            geo_level = "Municipality") |>
  filter(MunicipalityCode == "08221000")
```


Plot:

```{r}

```