---
title: 'Population data'
format: html
---


> Here I want to load population data, at Germany, Federal state and municipality level.

Libraries:

```{r}
library(here)
source(knitr::purl(here('ignore/libraries.qmd'), output = tempfile(), quiet = TRUE))
```


Credentials:

```{r}
gen_logincheck(database = "genesis")
gen_logincheck(database = "regio")
```

# 1. Create the dataset

## 1.1. Create the dataset: Germany

Code to create the dataset:
```{r}
pop.germany <- gen_table(
  name      = "12411-0001",
  database  = "genesis",
  language  = "en",
  startyear = 2010,
  endyear   = 2024
) |>
  select(
    country    = `1_variable_attribute_label`,
    year       = time,
    population = value
  ) |>
  mutate(
    date       = as.Date(paste0(substr(year, 1, 4), "-12-31")),
    population = suppressWarnings(as.numeric(population))
  ) |>
  select(country, date, population) |>
  arrange(date)
```

Code to saev the file here: 4. datasets\4.1. all germany\1. data

```{r}
arrow::write_parquet(
  pop.germany,
  here("4. datasets", "4.1. all germany", "1. data", "population.parquet")
)
```

Load the dataset:

```{r}
pop.germany <- arrow::read_parquet(
  here("4. datasets", "4.1. all germany", "1. data", "population.parquet")
)
```


---


## 1.2. Create the dataset: Federal state

Code to create the dataset:

```{r}
pop.federal_state <- gen_table(
  name      = "12411-0010",
  database  = "genesis",
  language  = "en",
  startyear = 2010,
  endyear   = 2024
)

federal_state.codes <- tibble::tibble(
  code = c("01","02","03","04","05","06","07","08","09","10","11","12","13","14","15","16"),
  nuts = c("DEF","DE6","DE9","DE5","DEA","DE7","DEB","DE1","DE2","DEC","DE3","DE4","DEM","DED","DEE","DEG")
)

pop.federal_state <- pop.federal_state |>
  select(
    code       = `1_variable_attribute_code`,
    name       = `1_variable_attribute_label`,
    year       = time,
    population = value
  ) |>
  mutate(
    date       = as.Date(paste0(substr(year, 1, 4), "-12-31")),
    population = suppressWarnings(as.numeric(population))
  ) |>
  select(-year) |>
  left_join(federal_state.codes, by = "code") |>
  select(FederalState.names = name, FederalStateCode = nuts, date, population) |>
  arrange(FederalStateCode, date)


```

Code to saev the file here: 4. datasets\4.2. federal state\1. data

```{r}
arrow::write_parquet(
  pop.federal_state,
  here("4. datasets", "4.2. federal state", "1. data", "population.parquet")
)
```

Load the dataset:

```{r}
pop.federal_state <- arrow::read_parquet(
  here("4. datasets", "4.2. federal state", "1. data", "population.parquet")
)
```


---

## 1.3. Create the dataset: Municipality


Code to create the dataset (yearly frequency):

```{r}
population_raw <- map_dfr(2008:2024, function(yr) {
    gen_table(
      name      = "12411-01-01-5",
      database  = "regio",
      language  = "en",
      startyear = yr,
      endyear   = yr
    )
  })


purrr::map(population_raw, function(i) {head(unique(i))}) |> kable()

# This is for select relevant variables
pop.municipality <- population_raw |>
  filter(is.na(`2_variable_attribute_code`)) |>
  select(
    MunicipalityCode = `1_variable_attribute_code`,
    Municipality     = `1_variable_attribute_label`,
    year             = time,
    population       = value
  ) |>
  mutate(
    date       = as.Date(paste0(substr(year, 1, 4), "-12-31")),
    population = suppressWarnings(as.numeric(population))
  ) |>
  select(Municipality, MunicipalityCode, date, population) |>
  arrange(Municipality, date)

```

Code to saev the file here: 4. datasets\4.2. federal state\1. data
I am not going to save just for Heidelberg, but for all municipalities. Once you load the dataset, you will have to filter for Heidelberg (08221000).

```{r}
arrow::write_parquet(
  pop.municipality,
  here("4. datasets", "4.3. Heidelberg", "1. data", "population.parquet")
)
```



Load the dataset:

```{r}
pop.municipality <- arrow::read_parquet(
  here("4. datasets", "4.3. Heidelberg", "1. data", "population.parquet")
)
```

---

# 2. Select 2023 population data
> Here I want to show how to produce the code to select 2023 population data for Germany, federal states and municipalities (Heidelberg). First, load the files, then filter them.


## 2.1. Germany
> Select 2023 population data for Germany. Load the file from "4. datasets\4.1. all germany\1. data" and filter it.

```{r}
pop.germany.2023 <- read_parquet(here("4. datasets", "4.1. all germany", "1. data", "population.parquet")) |>
  filter(date == as.Date("2023-12-31"))
```

## 2.2. Federal states
> Select 2023 population data for Federal states. Load the file from "4. datasets\4.2. federal state\1. data" and filter it.

```{r}
pop.federal_state.2023 <- read_parquet(here("4. datasets", "4.2. federal state", "1. data", "population.parquet")) |>
  filter(date == as.Date("2023-12-31"))
```

## 2.3. Municipalities (Heidelberg)
> Select 2023 population data for municipalities. Load the file from "44. datasets\4.3. Heidelberg\1. data" and filter it.

```{r}
pop.heidelberg.2023 <- read_parquet(here("4. datasets", "4.3. Heidelberg", "1. data", "population.parquet")) |>
  filter(date == as.Date("2023-12-31"), 
        code == "08221000")
```