---
title: 'Dataset for whole Germany'
format: html
---


> Here I want to create the dataset for Germany.

Libraries:

```{r}
library(here)
source(knitr::purl(here('ignore/libraries.qmd'), output = tempfile(), quiet = TRUE))
```


Libraries (necessary):

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  here,        # here()
  arrow,       # open_dataset(), read_parquet(), write_parquet(), collect()
  sfarrow,     # st_read_parquet()
  sf,          # st_drop_geometry()
  dplyr,       # rename, mutate, case_when, if_else, left_join, relocate,
               # filter, select, group_by, summarise, arrange, pull, ungroup, n
  tibble,      # tibble()
  tidyr,       # complete()
  lubridate,   # floor_date()
  stringr,     # str_pad()
  restatis     # gen_logincheck(), gen_table()
)
```



# 1. dataset

## 1.1. generate the dataset

Create the datasets:


```{r}
data <- open_dataset(here("01_data/EinheitenSolar_all.parquet"))


# Step 1: rename German → English (explicit mapping, column order differs from old dataset)
data <- data |>
  rename(
    UnitMastrNumber                          = EinheitMastrNummer,
    DateLastUpdate                           = DatumLetzteAktualisierung,
    LocationMastrNumber                      = LokationMaStRNummer,
    GridOperatorVerificationStatus           = NetzbetreiberpruefungStatus,
    GridOperatorVerificationDate             = NetzbetreiberpruefungDatum,
    PlantOperatorMastrNumber                 = AnlagenbetreiberMastrNummer,
    Country                                  = Land,
    FederalState                             = Bundesland,
    District                                 = Landkreis,
    Municipality                             = Gemeinde,
    MunicipalityCode                         = Gemeindeschluessel,
    PostalCode                               = Postleitzahl,
    City                                     = Ort,
    RegistrationDate                         = Registrierungsdatum,
    CommissioningDate                        = Inbetriebnahmedatum,
    UnitSystemStatus                         = EinheitSystemstatus,
    UnitOperationalStatus                    = EinheitBetriebsstatus,
    NotPresentInMigratedUnits                = NichtVorhandenInMigriertenEinheiten,
    PowerGenerationUnitName                  = NameStromerzeugungseinheit,
    Weic_NotAvailable                        = Weic_nv,
    PowerPlantNumber_NotAvailable            = Kraftwerksnummer_nv,
    EnergySource                             = Energietraeger,
    GrossCapacity                            = Bruttoleistung,
    NetNominalCapacity                       = Nettonennleistung,
    RemoteControllabilityGridOperator        = FernsteuerbarkeitNb,
    FeedInType                               = Einspeisungsart,
    AssignedActivePowerInverter              = ZugeordneteWirkleistungWechselrichter,
    NumberOfModules                          = AnzahlModule,
    TypeOfSolarSystem                        = ArtDerSolaranlage,
    PowerLimitation                          = Leistungsbegrenzung,
    UniformOrientationAndTiltAngle           = EinheitlicheAusrichtungUndNeigungswinkel,
    MainOrientation                          = Hauptausrichtung,
    MainOrientationTiltAngle                 = HauptausrichtungNeigungswinkel,
    UsageArea                                = Nutzungsbereich,
    EegMastrNumber                           = EegMaStRNummer,
    StorageAtSameLocation                    = SpeicherAmGleichenOrt,
    SecondaryOrientation                     = Nebenausrichtung,
    SecondaryOrientationTiltAngle            = NebenausrichtungNeigungswinkel,
    Street                                   = Strasse,
    StreetNotFound                           = StrasseNichtGefunden,
    HouseNumber_NotAvailable                 = Hausnummer_nv,
    HouseNumberNotFound                      = HausnummerNichtGefunden,
    Longitude                                = Laengengrad,
    Latitude                                 = Breitengrad,
    DateStartTemporaryDecommissioning        = DatumBeginnVoruebergehendeStilllegung,
    HouseNumber                              = Hausnummer,
    AddressAddition                          = Adresszusatz,
    RemoteControllabilityDirectMarketer      = FernsteuerbarkeitDv,
    PreviousPlantOperatorMastrNumber         = AltAnlagenbetreiberMastrNummer,  # new variable
    DateOfOperatorChange                     = DatumDesBetreiberwechsels,
    DateRegistrationOfOperatorChange         = DatumRegistrierungDesBetreiberwechsels,
    CadastralDistrict                        = Gemarkung,
    FieldParcelNumbers                       = FlurFlurstuecknummern,
    ConnectionToExtraHighOrHighVoltage       = AnschlussAnHoechstOderHochSpannung,
    DatePermanentDecommissioning             = DatumEndgueltigeStilllegung,
    SizeOfUsedAreaInHectares                 = GroesseDerInAnspruchGenommenenFlaecheInHektar,
    PreviousUsageTypeOfArea                  = VorherigerNutzungsartenbereichDerFlaeche,
    PredominantUsageTypeBeforeSolarConstruction = UeberwiegendeNutzungsartDerFlaecheVorErrichtung,
    NameOfSolarPark                          = NameDesSolarparks,
    AdditionalFeaturesOfAreaAndCurrentUsage  = ZusaetzlicheMerkmaleDerFlaecheUndDerAktuellenFlaechennutzung,
    GenMastrNumber                           = GenMastrNummer,
    CommissioningDateAtCurrentLocation       = InbetriebnahmedatumAmAktuellenStandort,
    DispatchResponsible                      = Einsatzverantwortlicher,
    CitizenEnergy                            = Buergerenergie,
    DateResumptionOfOperation                = DatumWiederaufnahmeBetrieb,
    PlannedCommissioningDate                 = GeplantesInbetriebnahmedatum,
    ClearanceHeight                          = LichteHoehe,
    PowerPlantNumber                         = Kraftwerksnummer
  )

# Step 2: convert numeric columns (XML parsing stores everything as character)
data <- data |>
  mutate(
    TypeOfSolarSystem           = as.integer(TypeOfSolarSystem),
    Country                     = as.integer(Country),
    FederalState                = as.integer(FederalState),
    GrossCapacity               = as.numeric(GrossCapacity),
    NetNominalCapacity          = as.numeric(NetNominalCapacity),
    AssignedActivePowerInverter = as.numeric(AssignedActivePowerInverter),
    NumberOfModules             = as.integer(NumberOfModules)
  )

# Step 3: lookup tables
C1 <- data.frame(
  TypeOfSolarSystem.names = c(
    "Structural_facilities", "Other_Structural_facilities", "Open_space",
    "Bodies_water", "Large_parking_lot", "Plug_in_sola_system", "Onshore&Offshore_wind_power"
  ),
  TypeOfSolarSystem = as.integer(c(853, 2484, 852, 3002, 3058, 2961, NA))
)

C2 <- tibble(
  Country = as.integer(c(84, 231, 90, 66, 198, 206, 106)),
  Country.sn = c("Germany", "Switzerland", "Denmark", "Belgium", "Netherlands", "Austria", "France")
)

C3 <- data.frame(
  FederalState = as.integer(c(
    1409, 1402, 1400, 1403, 1412, 1408, 1410, 1407,
    1414, 1411, 1413, 1405, 1415, 1406, 1401, 1404
  )),
  FederalState.names = c(
    "North Rhine-Westphalia", "Baden-Württemberg", "Brandenburg", "Bavaria",
    "Saarland", "Lower Saxony", "Rhineland-Palatinate", "Mecklenburg-Western Pomerania",
    "Saxony-Anhalt", "Schleswig-Holstein", "Saxony", "Hesse",
    "Thuringia", "Hamburg", "Berlin", "Bremen"
  ),
  FederalStateCode = c(
    "DEA", "DE1", "DE4", "DE2",
    "DEC", "DE9", "DEB", "DEM",
    "DEE", "DEF", "DED", "DE7",
    "DEG", "DE6", "DE3", "DE5"
  )
)

# district code from map (no geometry needed)
districts_map <- st_read_parquet(here("ignore/3. data/2. maps/districts.parquet")) |>
  st_drop_geometry() |>
  select(District, DistrictCode)

# Step 4: join lookups
data <- data |>
  left_join(C1, by = "TypeOfSolarSystem") |>
  relocate(TypeOfSolarSystem.names, .after = TypeOfSolarSystem) |>
  left_join(C2, by = "Country") |>
  relocate(Country.sn, .after = Country) |>
  left_join(C3, by = "FederalState") |>
  relocate(FederalState.names, .after = FederalState) |>
  relocate(FederalStateCode, .after = FederalState.names) |>
  left_join(districts_map, by = "District") |>
  relocate(DistrictCode, .after = District)

# Step 5: solar_category
reform_date <- as.Date("2024-05-16")

data <- data |>
  mutate(
    RegistrationDate  = as.Date(RegistrationDate),
    CommissioningDate = as.Date(CommissioningDate),
    solar_category = case_when(
      TypeOfSolarSystem == 2961 & RegistrationDate <  reform_date & AssignedActivePowerInverter <= 0.6 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 & RegistrationDate >= reform_date & AssignedActivePowerInverter <= 0.8 & GrossCapacity <= 2.0 ~ "Plug_in_sola_system - Balcony",
      TypeOfSolarSystem == 2961 ~ "Plug_in_sola_system - Non-Balcony",
      TRUE ~ TypeOfSolarSystem.names
    )
  )
# Step 6: filter Germany
data <- data |>
  filter(Country.sn == "Germany")

# Step 7: date variables (week, month, quarter, year)
data <- data |>
  mutate(
    date.registration.day     = RegistrationDate,
    date.registration.week    = floor_date(RegistrationDate,  "week"),
    date.registration.month   = floor_date(RegistrationDate,  "month"),
    date.registration.quarter = floor_date(RegistrationDate,  "quarter"),
    date.registration.year    = floor_date(RegistrationDate,  "year"),
    date.commission.day       = CommissioningDate,
    date.commission.week      = floor_date(CommissioningDate, "week"),
    date.commission.month     = floor_date(CommissioningDate, "month"),
    date.commission.quarter   = floor_date(CommissioningDate, "quarter"),
    date.commission.year      = floor_date(CommissioningDate, "year")
  )

# Step 8: fix MunicipalityCode
data <- data |>
  mutate(MunicipalityCode = str_pad(as.character(MunicipalityCode), 8, pad = "0"))
```

Collapse every category other than `Balcony PVs` into a single `Non-balcony PVs` group:

```{r}
data <- data |>
  mutate(solar_category = if_else(solar_category == "Plug_in_sola_system - Balcony",
                                  "Balcony PVs",
                                  "Non-balcony PVs"))
```


Frequency:

```{r}
germany_day <- data |>
  group_by(solar_category, date.registration.day) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.day = seq(min(date.registration.day), max(date.registration.day), by = "day"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.day) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_week <- data |>
  group_by(solar_category, date.registration.week) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.week = seq(min(date.registration.week), max(date.registration.week), by = "week"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.week) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_month <- data |>
  group_by(solar_category, date.registration.month) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.month = seq(min(date.registration.month), max(date.registration.month), by = "month"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.month) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_quarter <- data |>
  group_by(solar_category, date.registration.quarter) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.quarter = seq(min(date.registration.quarter), max(date.registration.quarter), by = "quarter"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.quarter) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()

germany_year <- data |>
  group_by(solar_category, date.registration.year) |> summarise(n = n(), .groups = "drop") |> collect() |>
  complete(solar_category, date.registration.year = seq(min(date.registration.year), max(date.registration.year), by = "year"),
           fill = list(n = 0)) |>
  arrange(solar_category, date.registration.year) |>
  group_by(solar_category) |> mutate(n_cumulative = cumsum(n)) |> ungroup()
```






## 1.2. Load te datasets
> Here I want to load the datasets.

Code to laod datasets:

```{r}

```



---

# 2. population data
> here I want to show how to load the population data for the whole Germany.

## 2.1. Generate the dataset
> Here I want to show hwo to create the daatset from zero. Below you find the version where you just need to load it.

Credentials:

```{r}
gen_logincheck(database = "genesis")
gen_logincheck(database = "regio")
```

Code to generate the dataset:

```{r}
pop.germany <- gen_table(
  name      = "12411-0001",
  database  = "genesis",
  language  = "en",
  startyear = 2010,
  endyear   = 2024
) |>
  select(
    country    = `1_variable_attribute_label`,
    year       = time,
    population = value
  ) |>
  mutate(
    date       = as.Date(paste0(substr(year, 1, 4), "-12-31")),
    population = suppressWarnings(as.numeric(population))
  ) |>
  select(country, date, population) |>
  arrange(date)
```

Filter the dataste for 2023 population data:

```{r}
pop.germany.2023 <- pop.germany |>
  filter(date == as.Date("2023-12-31"))
```


## 2.2. Load the dataset
> Here I want to load the dataset I have already created.

Load the file:

```{r}
pop.germany.2023 <- read_parquet(here("4. datasets", "4.1. all germany", "1. data", "population.parquet")) |>
  filter(date == as.Date("2023-12-31"))
```

---

# 3. Merge

## 3.1. Generate the dataset
> Here I want to show how to create the dataset.

Code to include the population data in the dataset:

```{r}
pop_2023 <- pop.germany.2023 |> pull(population)
```


Code to add the variables:

```{r}
germany_day <- germany_day |>
  mutate(
    n_per100k            = (n            / pop_2023) * 100000,
    n_cumulative_per100k = (n_cumulative / pop_2023) * 100000
  )

germany_week <- germany_week |>
  mutate(
    n_per100k            = (n            / pop_2023) * 100000,
    n_cumulative_per100k = (n_cumulative / pop_2023) * 100000
  )

germany_month <- germany_month |>
  mutate(
    n_per100k            = (n            / pop_2023) * 100000,
    n_cumulative_per100k = (n_cumulative / pop_2023) * 100000
  )

germany_quarter <- germany_quarter |>
  mutate(
    n_per100k            = (n            / pop_2023) * 100000,
    n_cumulative_per100k = (n_cumulative / pop_2023) * 100000
  )

germany_year <- germany_year |>
  mutate(
    n_per100k            = (n            / pop_2023) * 100000,
    n_cumulative_per100k = (n_cumulative / pop_2023) * 100000
  )
```

## 3.2. Load the dataset

Code to save them and load:

Save the datasets in parquet format:

```{r}
out_dir <- here("4. datasets", "4.1. all germany", "1. data")

write_parquet(germany_day,     file.path(out_dir, "germany_day.parquet"))
write_parquet(germany_week,    file.path(out_dir, "germany_week.parquet"))
write_parquet(germany_month,   file.path(out_dir, "germany_month.parquet"))
write_parquet(germany_quarter, file.path(out_dir, "germany_quarter.parquet"))
write_parquet(germany_year,    file.path(out_dir, "germany_year.parquet"))
```

Load the datasets:

```{r}
out_dir <- here("4. datasets", "4.1. all germany", "1. data")

germany_day     <- read_parquet(file.path(out_dir, "germany_day.parquet"))
germany_week    <- read_parquet(file.path(out_dir, "germany_week.parquet"))
germany_month   <- read_parquet(file.path(out_dir, "germany_month.parquet"))
germany_quarter <- read_parquet(file.path(out_dir, "germany_quarter.parquet"))
germany_year    <- read_parquet(file.path(out_dir, "germany_year.parquet"))
```